---
format:
  dashboard:
    orientation: columns
    scrolling: true
---
```{r}
#| message: FALSE
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")
TheList <- c("5L")

# Updating Data
# Data processing (parsing new QC .fcs/.csv files into the Archive) is handled
# by TheScript_5L.R, which runs on the instrument computer. This dashboard only
# *displays* the already-processed Archive data, so the parsing calls below are
# left commented out to keep the render reproducible and side-effect free.
# walk(.x=TheList, MainFolder=MainFolder, .f=Luciernaga:::DailyQCParse)
# walk(.x=TheList, .f=Luciernaga:::QCBeadParse, MainFolder=MainFolder)
```

```{r}
Gain_5L <- Luciernaga:::CurrentData(x="5L", MainFolder=MainFolder, type = "Gain")
TheDate <- Gain_5L %>% slice(1) %>% pull(DateTime) %>% as.Date()
```

```{r}
WindowOfInterest <- Sys.time() - months(12)

Gain_5L <- Gain_5L %>% filter(DateTime >= WindowOfInterest)
```

```{r}
Data <- read.csv("data/reference/AuroraMaintenance.csv", check.names=FALSE)

Data <- Data %>% filter(!str_detect(reason, "lean"))

Repair5L <- Data %>% filter(instrument %in% "5L")
```

```{r}
x <- Gain_5L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "Gain")]

UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)
LaserGains <- TheColumns[str_detect(TheColumns, "Laser")]
LaserDelayGains <- LaserGains[str_detect(LaserGains, "Delay")]
LaserDelayGains <- Luciernaga:::ColorPriority(LaserDelayGains)
LaserPowerGains <- LaserGains[str_detect(LaserGains, "Power")]
LaserPowerGains <- Luciernaga:::ColorPriority(LaserPowerGains)
ScalingGains <- TheColumns[str_detect(TheColumns, "Scaling")]
ScalingGains <- Luciernaga:::ColorPriority(ScalingGains)

UltraVioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair5L)

VioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict = TRUE, YAxisLabel = "Gain",
                      RepairVisits=Repair5L)

BluePlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair5L)

YellowGreenPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair5L)

RedPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                     RepairVisits=Repair5L)

ScatterPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair5L)

LaserDelayPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserDelayGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair5L)

LaserPowerPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserPowerGains,
                                plotType = "individual", returntype = "plots",
                                YAxisLabel = " ", RepairVisits=Repair5L)

ScalingPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScalingGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair5L)
```

```{r}
x <- Gain_5L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")

TheColumnNames <- TheColumns[str_detect(TheColumns, "rCV")]
UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

UltraVioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair5L)

VioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict=TRUE, YAxisLabel = "%rCV",
                      RepairVisits=Repair5L)

BluePlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair5L)

YellowGreenPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair5L)

RedPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                     RepairVisits=Repair5L)

ScatterPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair5L)
```

## {.sidebar}
Dashboard data for the **Cytek Aurora 5L** last updated on **`r TheDate`**

**First Column: Gain** Gain (Voltage) values set for instrument after QC. Changes over time reflective of laser health.
**Second Column: rCV** Percentage Robust Coefficient of Variation (rCV) after QC. Higher values reflect decreased resolution between positive and negative for that detector.

For additional information concerning individual parameter tabs, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek cytometers at the Sydney Cytometry Core Facility

For additional flow cytometry resources, visit our [website](https://sydneycytometry.org.au/)

This dashboard was created with [Quarto](https://quarto.org/), based on work by [David Rach](https://github.com/DavidRach).

## Gain {.tabset}

```{r}
#| title: UltraViolet
htmltools::tagList(lapply(UltraVioletPlotsGain, ggplotly))
```

```{r}
#| title: Violet
htmltools::tagList(lapply(VioletPlotsGain, ggplotly))
```

```{r}
#| title: Blue
htmltools::tagList(lapply(BluePlotsGain, ggplotly))
```

```{r}
#| title: YellowGreen
htmltools::tagList(lapply(YellowGreenPlotsGain, ggplotly))
```

```{r}
#| title: Red
htmltools::tagList(lapply(RedPlotsGain, ggplotly))
```

```{r}
#| title: Scatter
htmltools::tagList(lapply(ScatterPlotsGain, ggplotly))
```

```{r}
#| title: LaserPower
htmltools::tagList(lapply(LaserPowerPlotsGain, ggplotly))
```

```{r}
#| title: LaserDelay
htmltools::tagList(lapply(LaserDelayPlotsGain, ggplotly))
```

```{r}
#| title: LaserScaling
htmltools::tagList(lapply(ScalingPlotsGain, ggplotly))
```

## rCV {.tabset}

```{r}
#| title: UltraViolet
htmltools::tagList(lapply(UltraVioletPlotsRCV, ggplotly))
```

```{r}
#| title: Violet
htmltools::tagList(lapply(VioletPlotsRCV, ggplotly))
```

```{r}
#| title: Blue
htmltools::tagList(lapply(BluePlotsRCV, ggplotly))
```

```{r}
#| title: YellowGreen
htmltools::tagList(lapply(YellowGreenPlotsRCV, ggplotly))
```

```{r}
#| title: Red
htmltools::tagList(lapply(RedPlotsRCV, ggplotly))
```

```{r}
#| title: Scatter
htmltools::tagList(lapply(ScatterPlotsRCV, ggplotly))
```

```{r}
#| title: LaserPower
htmltools::tagList(lapply(LaserPowerPlotsGain, ggplotly))
```

```{r}
#| title: LaserDelay
htmltools::tagList(lapply(LaserDelayPlotsGain, ggplotly))
```

```{r}
#| title: LaserScaling
htmltools::tagList(lapply(ScalingPlotsGain, ggplotly))
```
