---
title: "6 Peak Ultra Bead Tracking — Aurora EVO (5L)"
subtitle: "Spectral QC: instrument-performance tracking from URCP-38-2K beads"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    fig-width: 9
---

```{r}
#| label: load
#| include: false
library(ggcyto)   # scale_*_flowjo_biexp; pulls in flowCore (masks dplyr verbs)
library(ggplot2)
library(tidyr)
library(readr)
library(plotly)
library(dplyr)    # LAST so dplyr verbs win the namespace masking

results_file <- if (file.exists("6peak Ultra_tracking_spectral/peak_metrics.csv")) {
  "6peak Ultra_tracking_spectral/peak_metrics.csv"
} else "peak_metrics.csv"
pm <- read_csv(results_file, show_col_types = FALSE)
pm$datetime <- as.POSIXct(pm$datetime, tz = "UTC")
pm$date     <- as.Date(pm$datetime)

# URCP-38-2K resolves into 6 populations: blank particles (peak 1) + 5 fluorescent
# rainbow levels (peaks 2-6). Peak 6 is the brightest.
n_peaks  <- 6
biex_max <- 4194304; biex_pos <- 5.62; biex_neg <- 0; biex_width_basis <- -250

# Detector acquisition order (UV -> V -> B -> YG -> R) from the archive.
laser_levels <- c("UV", "V", "B", "YG", "R")
laser_of <- function(d) sub("[0-9]+-A$", "", as.character(d))
det_order <- pm |> distinct(detector) |>
  mutate(laser = laser_of(detector),
         num   = as.integer(sub("-A$", "", sub("^(UV|V|B|YG|R)", "", detector)))) |>
  arrange(match(laser, laser_levels), num)
detector_channels <- det_order$detector
pm$detector <- factor(pm$detector, levels = detector_channels)

# Run axis. Unlike the 8-peak set (one run/day) URCP can be acquired more than
# once a day, so the tracking unit is the acquisition TIMESTAMP. run_label is an
# ORDERED factor (chronological) used as the categorical x on every panel; it
# behaves like `date` did on the 8-peak page (arrange/first/last/min/max all work).
run_ord  <- pm |> distinct(run, datetime) |> arrange(datetime)
run_levs <- format(run_ord$datetime, "%Y-%m-%d %H:%M")
pm$run_label <- factor(format(pm$datetime, "%Y-%m-%d %H:%M"), levels = run_levs, ordered = TRUE)
runs   <- levels(pm$run_label); n_runs <- length(runs)
base_run <- runs[1]

# ---- GAIN cross-reference: the daily QC gain applied on each run ------------
# On the Aurora, daily QC pins bead MFI to a target by adjusting APD gain, so raw
# MFI is largely held constant by design. Dividing MFI by gain de-pins it, giving
# a ~gain-independent signal; CV-based sensitivity (Q/B) is gain-invariant. Gain
# is recorded once per day, so it is joined on the calendar date.
gfile <- "data/EVO/Archive/ArchivedDataEVO.csv"
has_gain <- file.exists(gfile)
if (has_gain) {
  g <- read.csv(gfile, check.names = FALSE)
  g$Date <- as.Date(lubridate::ymd_hms(g$DateTime))
  glong <- g |>
    select(Date, ends_with("-Gain")) |>
    pivot_longer(-Date, names_to = "detg", values_to = "gain") |>
    mutate(detector = paste0(sub("-Gain$", "", detg), "-A")) |>
    group_by(Date, detector) |> summarise(gain = dplyr::first(gain), .groups = "drop")
  pm <- pm |> left_join(glong, by = c("date" = "Date", "detector"))
} else pm$gain <- NA_real_
# Gain de-pins MFI. When SOME runs have recorded gain but a newer run does NOT yet
# (its daily-QC gain has not been exported to the archive), do NOT substitute raw
# MFI for that run: mixing raw and gain-normalised values across runs makes a
# spurious order-of-magnitude jump on the gain panels. Instead drop the gainless
# run from gain-normalised views (mfi_norm = NA). Only when NO run has any gain do
# we fall back to raw MFI globally (a consistent, if relative, scale).
any_gain <- has_gain && any(is.finite(pm$gain))
pm <- pm |> mutate(mfi_norm = dplyr::case_when(
  !is.na(gain) & gain > 0 ~ median / gain,
  any_gain                ~ NA_real_,
  TRUE                    ~ median))
has_gain <- any_gain
runs_no_gain <- pm |> filter(peak == n_peaks) |>
  group_by(run_label) |> summarise(g = any(is.finite(gain)), .groups = "drop") |>
  filter(!g) |> pull(run_label) |> as.character()

# ---- per-detector x per-run metrics ----
# Sensitivity (Q & B): CV^2 = intrinsic_CV^2 + (1/Q)*(1/x) + (B/Q)*(1/x^2),
# where x is the per-event intensity. Two flavours of x:
#   - xvar = "mfi_norm" (default): gain-normalised MFI -> Q/B in arbitrary but
#     gain-stabilised units; available for all 64 detectors (relative drift).
#   - xvar = "MEF": URCP-38-2K assigned MEF -> Q/B in true MEF units (STN-17),
#     gain-independent; anchor detectors only (absolute sensitivity).
# GUARD: Q = 1/slope explodes when the slope -> 0, so a noisy run can emit a phantom
# Q spike. We suppress Q/B (-> NA, with a `qflag` reason) when the fit is ill-
# conditioned: near-flat/negative slope, poor R^2, a non-monotonic CV profile
# (CV^2 should rise with 1/x), or an inflated brightest-peak CV.
qb_fit <- function(d, xvar = "mfi_norm", true_b = FALSE,
                   min_r2 = 0.8, min_mono = 0.3, max_bright_cv = 15) {
  d <- d[which(d$resolved & d$in_range & !is.na(d[[xvar]]) & d[[xvar]] > 0), ]
  if (nrow(d) < 3)
    return(tibble(intr_cv = NA_real_, Q = NA_real_, B = NA_real_, npts = nrow(d), qflag = "n<3"))
  cv2 <- (d$rcv / 100)^2; inv <- 1 / d[[xvar]]
  f2  <- lm(cv2 ~ inv); s1 <- coef(f2)[2]
  r2  <- summary(f2)$r.squared
  mono <- suppressWarnings(stats::cor(inv, cv2, method = "spearman"))   # CV^2 vs 1/x: want > 0
  bcv  <- d$rcv[which.max(d[[xvar]])]                                   # brightest in-range peak rCV
  intr <- sqrt(pmax(coef(f2)[1], 0))
  flags <- c(if (!is.finite(s1) || s1 <= 0)              "flat/neg slope"          else NULL,
             if (is.finite(r2)   && r2   < min_r2)        sprintf("R2=%.2f", r2)    else NULL,
             if (is.finite(mono) && mono < min_mono)      "non-monotonic"          else NULL,
             if (is.finite(bcv)  && bcv  > max_bright_cv) sprintf("brightCV=%.0f%%", bcv) else NULL)
  if (length(flags))
    return(tibble(intr_cv = intr, Q = NA_real_, B = NA_real_, npts = nrow(d),
                  qflag = paste(flags, collapse = "; ")))
  s2 <- if (nrow(d) >= 4) coef(lm(cv2 ~ inv + I(inv^2)))[3] else NA_real_
  B  <- if (true_b) s2 / s1 else s2
  tibble(intr_cv = intr, Q = 1 / s1, B = B, npts = nrow(d), qflag = NA_character_)
}
qb <- pm |> group_by(run_label, detector) |> group_modify(~ qb_fit(.x)) |> ungroup()

# Resolution, brightest-/blank-peak %CV, brightest gain-normalised intensity
res   <- pm |> group_by(run_label, detector) |> summarise(n_resolved = sum(resolved), .groups = "drop")
cvtop <- pm |> filter(peak == n_peaks, in_range) |> select(run_label, detector, rcv) |> rename(cvt = rcv)
cvp1  <- pm |> filter(peak == 1, in_range) |> mutate(sd1 = rcv / 100 * median) |>
  select(run_label, detector, sd1)   # blank NOISE FLOOR = absolute robust SD; %CV unstable near 0 mean
ptop  <- pm |> filter(peak == n_peaks, in_range) |> select(run_label, detector, median, gain, mfi_norm)

# Staining index (lowest resolved peak vs blank peak 1)
si_one <- function(g) {
  cand <- g$peak[g$peak > 1 & g$resolved & g$in_range]
  if (!length(cand)) return(tibble(si = NA_real_, pos_peak = NA_integer_))
  p <- min(cand); m1 <- g$median[g$peak == 1]; rsd1 <- abs(g$rcv[g$peak == 1] * m1 / 100)
  tibble(si = (g$median[g$peak == p] - m1) / (2 * rsd1), pos_peak = as.integer(p))
}
si <- pm |> group_by(run_label, detector) |> group_modify(~ si_one(.x)) |> ungroup()

# Linearity (rank-based, all 64): measured MFI vs peak rank, per detector-run.
lin <- pm |> filter(resolved, in_range) |> group_by(run_label, detector) |>
  filter(n() >= 3) |>
  summarise(r2 = summary(lm(log10(median) ~ peak))$r.squared, .groups = "drop")

# ---- MEF calibration (SPHERO URCP-38-2K certificate) for anchor detectors ---
# fluor -> detector map (standard Cytek Aurora 5L peak channels; every pairing was
# validated against these beads by a log-log median-vs-MEF fit, R^2 >= 0.96). The
# two blue-laser cross-excited tandems (PE-TexasRed/PE-Cy5 at 488ex) are excluded.
mef_map <- tibble::tribble(
  ~fluor,              ~detector,
  "BUV395",            "UV2-A",
  "BUV737",            "UV14-A",
  "BV421",             "V1-A",
  "BV510",             "V5-A",
  "BV605",             "V10-A",
  "BV650",             "V11-A",
  "BV711",             "V13-A",
  "BV786",             "V15-A",
  "FITC",              "B2-A",
  "PerCP-Cy5.5",       "B9-A",
  "PE",                "YG1-A",
  "PE-TexasRed(561)",  "YG3-A",
  "PE-Cy5(561)",       "YG5-A",
  "PE-Cy5.5",          "YG6-A",
  "PE-Cy7",            "YG9-A",
  "APC",               "R1-A",
  "Alexa700",          "R4-A",
  "APC-Cy7",           "R7-A")
mef_file <- "data/reference/URCP-38-2K_MEF.csv"
has_mef  <- file.exists(mef_file)
if (has_mef) {
  mefv <- readr::read_csv(mef_file, show_col_types = FALSE) |> inner_join(mef_map, by = "fluor")
  # PEAK ALIGNMENT: data peak 1 = blank particles (no assigned MEF; the
  # certificate's peak 1 is also blank/NA and absent from the CSV). The 5
  # fluorescent levels are data peaks 2-6 and line up DIRECTLY with certificate
  # peaks 2-6, so the join is 1:1 on peak (median R^2 ~0.9999, slope ~1.17).
  lin_mef <- pm |> mutate(detector = as.character(detector)) |>
    inner_join(mefv, by = c("detector", "peak")) |>
    filter(resolved, in_range, median > 0, MEF > 0)
  # log-log calibration per anchor x run (slope ~1 if linear; intercept holds gain)
  calib <- lin_mef |> group_by(run_label, detector, fluor) |> filter(n() >= 3) |>
    summarise(b = coef(lm(log10(median) ~ log10(MEF)))[2],
              a = coef(lm(log10(median) ~ log10(MEF)))[1],
              r2 = summary(lm(log10(median) ~ log10(MEF)))$r.squared, .groups = "drop")
  # Detection limit in MEF: blank (peak 1) + 2 robust SD, inverted through the
  # calibration. The applied gain cancels, so MEF_LOD is gain-independent.
  blank <- pm |> filter(peak == 1) |> mutate(detector = as.character(detector)) |>
    semi_join(mef_map, by = "detector") |>
    transmute(run_label, detector, thr = median * (1 + 2 * rcv / 100))
  sens <- calib |> left_join(blank, by = c("run_label", "detector")) |>
    mutate(MEF_LOD = 10^((log10(thr) - a) / b))
  # MEF-unit Q & B (STN-17): same CV^2-vs-1/x regression as `qb`, x = assigned MEF
  # -> Q/B in true molecules-of-equivalent-fluorophore units, gain-independent.
  qb_mef <- lin_mef |> group_by(run_label, detector, fluor) |>
    group_modify(~ qb_fit(.x, xvar = "MEF", true_b = TRUE)) |> ungroup()
  # ---- Literal STN-17 closed form (independent cross-check) ----
  # Blank (peak 1) extrapolated through the calibration; dim reference bead = the
  # lowest data peak carrying an assigned MEF (peak 2); brightest = peak 6
  # (high-signal CV floor).
  dim_ref_peak <- if (nrow(lin_mef)) min(lin_mef$peak) else 2L
  p_cv <- function(p) pm |> filter(peak == p, median > 0) |>
    mutate(detector = as.character(detector)) |> semi_join(mef_map, by = "detector") |>
    transmute(run_label, detector, m = median, cv = rcv / 100)
  b1   <- p_cv(1)            |> rename(m1 = m, cv1 = cv)
  bdim <- p_cv(dim_ref_peak) |> rename(mdim = m, cvdim = cv) |>
    left_join(mefv |> filter(peak == dim_ref_peak) |> select(detector, MEFdim = MEF), by = "detector")
  btop <- p_cv(n_peaks) |> transmute(run_label, detector, cvtop = cv)
  stn17 <- calib |> select(run_label, detector, fluor, a, b) |>
    left_join(b1,   by = c("run_label", "detector")) |>
    left_join(bdim, by = c("run_label", "detector")) |>
    left_join(btop, by = c("run_label", "detector")) |>
    mutate(
      MEF_blank = 10^((log10(m1) - a) / b),                  # blank MEF via calib
      Q_stn     = 1 / (cv1^2 * MEF_blank),
      SD2_blank = MEF_blank^2 * (cv1^2   - cvtop^2),
      SD2_dim   = MEFdim^2    * (cvdim^2 - cvtop^2),
      B_stn     = (SD2_blank / SD2_dim) * MEF_blank)
}

# Spectral signature drift: cosine similarity of the 64-detector signature of each
# peak vs the baseline run (gain-normalised). 1 = unchanged; drops = optical drift.
cos_sim <- function(a, b) sum(a * b) / (sqrt(sum(a * a)) * sqrt(sum(b * b)))
sig <- pm |> filter(in_range) |> select(run_label, peak, detector, mfi_norm) |>
  arrange(peak, detector)
spec <- sig |> group_by(peak) |> group_modify(function(g, k) {
  bv <- g |> filter(run_label == base_run)
  g |> group_by(run_label) |> summarise(
    cosine = cos_sim(mfi_norm, bv$mfi_norm[match(detector, bv$detector)]), .groups = "drop")
}) |> ungroup()

# Combined per-detector-run table + baseline deltas (for the overview heatmap)
metrics <- res |>
  left_join(qb,    by = c("run_label","detector")) |>
  left_join(cvtop, by = c("run_label","detector")) |>
  left_join(si,    by = c("run_label","detector")) |>
  left_join(ptop |> select(run_label, detector, mfi_norm), by = c("run_label","detector"))
add_base <- function(df, v) df |> group_by(detector) |> arrange(run_label) |>
  mutate("{v}_d" := (.data[[v]] - first(.data[[v]])) / abs(first(.data[[v]])) * 100) |> ungroup()
for (v in c("Q","B","cvt","si","mfi_norm","n_resolved")) metrics <- add_base(metrics, v)

# Control limits (mean +/- 2 SD) across runs, per detector, for the trend bands.
clim <- function(df, v) df |> group_by(detector) |>
  summarise(mu = mean(.data[[v]], na.rm = TRUE), sd = sd(.data[[v]], na.rm = TRUE), .groups = "drop") |>
  mutate(lcl = mu - 2*sd, ucl = mu + 2*sd)
cv_y_scale <- scale_y_continuous(expand = expansion(mult = c(0.06, 0.18)))

# Helper: per-detector x run heatmap (rows = detector in acquisition order)
heat <- function(df, fill, title, sub = NULL, diverging = FALSE, limitsym = FALSE) {
  p <- ggplot(df, aes(run_label, detector, fill = .data[[fill]],
                      text = sprintf("%s | %s<br>%s: %.3g", detector, run_label, fill, .data[[fill]]))) +
    geom_tile(colour = "white", linewidth = 0.1) +
    scale_y_discrete(limits = rev(detector_channels)) +
    labs(x = NULL, y = NULL, title = title, subtitle = sub, fill = "") +
    theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                    axis.text.x = element_text(angle = 30, hjust = 1),
                                    plot.subtitle = element_text(size = 8, colour = "grey40"))
  if (diverging) {
    lim <- if (limitsym) max(abs(df[[fill]]), na.rm = TRUE) else NA
    p <- p + scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0,
                                  limits = if (is.na(lim)) NULL else c(-lim, lim))
  } else p <- p + scale_fill_viridis_c(option = "viridis")
  p
}
gp <- function(p, h = 620) ggplotly(p, tooltip = "text", height = h) |>
  plotly::config(displayModeBar = FALSE)

# ---- Per-detector fold-change reference for the FC maps --------------------
# Absolute heatmaps span orders of magnitude across detectors, so the companion
# "FC" maps show each detector relative to its OWN reference. Two modes:
#   * NULL  -> DYNAMIC / preliminary: reference = each detector's own best run so far.
#   * c("YYYY-MM-DD HH:MM","...") -> LOCKED: per-detector robust centre (median) + SD
#     over a qualification window, frozen to fc_ref_file (two-sided deviation view).
fc_ref_window <- NULL                                   # <- set a run-label range to LOCK
fc_ref_file   <- "data/reference/EVO_6peakUltra_reference.csv"
fc_locked     <- !is.null(fc_ref_window)
fc_ref        <- NULL
if (fc_locked) {
  if (file.exists(fc_ref_file)) {
    fc_ref <- readr::read_csv(fc_ref_file, show_col_types = FALSE)
  } else {
    fc_metrics <- bind_rows(
      qb    |> transmute(run_label, detector, metric = "Q",        value = Q),
      si    |> transmute(run_label, detector, metric = "si",       value = si),
      cvp1  |> transmute(run_label, detector, metric = "sd1",      value = sd1),
      ptop  |> transmute(run_label, detector, metric = "mfi_norm", value = mfi_norm))
    fc_ref <- fc_metrics |>
      filter(run_label >= fc_ref_window[1], run_label <= fc_ref_window[2], is.finite(value)) |>
      group_by(detector, metric) |>
      summarise(ref = median(value), ref_sd = sd(value), n_ref = dplyr::n(), .groups = "drop")
    readr::write_csv(fc_ref, fc_ref_file)
  }
}

# DYNAMIC reference: each detector vs its own best (one-sided). deg >= 0 (0 = best).
best_fc <- function(df, v, higher_better = TRUE) {
  df |> group_by(detector) |> mutate(
    .val  = ifelse(is.finite(.data[[v]]) & .data[[v]] > 0, .data[[v]], NA_real_),  # log2 needs > 0
    .best = { x <- .val[is.finite(.val)]
              if (!length(x)) NA_real_ else if (higher_better) max(x) else min(x) },
    deg   = if (higher_better) log2(.best / .val) else log2(.val / .best),
    fold  = 2^deg) |> ungroup()
}
heat_fc <- function(df, v, title, sub = NULL, higher_better = TRUE) {
  if (fc_locked) {
    r <- fc_ref |> filter(metric == v) |> transmute(detector = as.character(detector), ref, ref_sd)
    d <- df |> mutate(detector = as.character(detector)) |> left_join(r, by = "detector") |>
      mutate(value = .data[[v]],
             lfc   = log2(value / ref)        * (if (higher_better) -1 else 1),
             sddev = (value - ref) / ref_sd   * (if (higher_better) -1 else 1))
    hi <- stats::quantile(abs(d$lfc), 0.95, na.rm = TRUE); if (!is.finite(hi) || hi <= 0) hi <- 1
    ggplot(d, aes(run_label, detector, fill = pmax(pmin(lfc, hi), -hi),
          text = sprintf("%s | %s<br>%s = %.3g<br>%+.2f log₂ vs normal | %+.1f SD",
                         detector, run_label, v, value, lfc, sddev))) +
      geom_tile(colour = "white", linewidth = 0.1) +
      scale_y_discrete(limits = rev(detector_channels)) +
      scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b", midpoint = 0,
                           limits = c(-hi, hi), name = "log₂ vs\nqualified\nnormal") +
      labs(x = NULL, y = NULL, title = title,
           subtitle = "deviation from locked reference (red = worse)") +
      theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                      axis.text.x = element_text(angle = 30, hjust = 1),
                                      plot.subtitle = element_text(size = 8, colour = "grey40"))
  } else {
    d  <- best_fc(df, v, higher_better)
    hi <- stats::quantile(d$deg, 0.95, na.rm = TRUE); if (!is.finite(hi) || hi <= 0) hi <- 1
    ggplot(d, aes(run_label, detector, fill = pmin(deg, hi),
          text = sprintf("%s | %s<br>%s = %.3g<br>%.2f× its own best", detector, run_label, v, .data[[v]], fold))) +
      geom_tile(colour = "white", linewidth = 0.1) +
      scale_y_discrete(limits = rev(detector_channels)) +
      scale_fill_gradient(low = "white", high = "#b2182b", limits = c(0, hi),
                          name = "log₂ fold\nvs best", na.value = "grey92") +
      labs(x = NULL, y = NULL, title = title,
           subtitle = if (is.null(sub)) "preliminary: vs each detector's best so far" else sub) +
      theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                      axis.text.x = element_text(angle = 30, hjust = 1),
                                      plot.subtitle = element_text(size = 8, colour = "grey40"))
  }
}
```

Secondary QC for the Cytek Aurora EVO from SPHERO **URCP-38-2K** Ultra Rainbow
beads. The set resolves into **6 populations** — blank particles (peak 1) plus 5
fluorescent rainbow levels (peaks 2–6). **Note on the Aurora:** daily QC adjusts
each APD's *gain* to pin bead MFI to a target, so raw MFI is held ~constant by
design — the informative signals are the ones gain can't compensate:
**sensitivity (Q & B)**, **resolution**, **%CV**, **linearity**, and
**spectral-signature stability**. Heavy FCS processing is done once by
`scripts/Process6PeakUltra.R`; this page reads the `peak_metrics.csv` archive
(+ the daily-QC gain) so it stays light.

```{r}
#| label: prelim-note
#| echo: false
#| output: asis
if (n_runs < 5) cat(sprintf(
  "::: {.callout-warning appearance='simple'}\nOnly **%d run(s)** so far — trends and the mean ± 2 SD control limits are **preliminary** and will firm up as more runs accumulate.\n:::\n\n", n_runs))
```

# Executive summary

```{r}
#| label: run-status
#| results: asis
status <- pm |> filter(peak == n_peaks) |> group_by(run_label) |>
  summarise(saturated = sum(!in_range), total = n(), .groups = "drop") |> arrange(run_label)
for (i in seq_len(nrow(status))) { r <- status[i, ]
  cat(sprintf("::: {.%s}\n**Run %s** — %d / %d detectors saturated at peak 6 (brightest).\n:::\n\n",
              if (r$saturated == 0) "callout-note" else "callout-warning",
              as.character(r$run_label), r$saturated, r$total)) }
```

Latest run vs the baseline run (**`r base_run`**), per detector. Blue = improved,
red = degraded (sign adjusted so red is always *worse*). Hover for values.

```{r}
#| label: overview-heat
latest <- runs[length(runs)]
ov <- metrics |> filter(run_label == latest) |>
  transmute(detector,
            `Sensitivity Q` = Q_d,                 # higher = better -> red when negative
            `Background B`  = -B_d,                 # higher B worse -> flip
            `Peak-6 %CV`    = -cvt_d,               # higher CV worse -> flip
            `Staining idx`  = si_d,
            `Peaks resolved`= n_resolved_d,
            `MFI/gain`      = mfi_norm_d) |>
  pivot_longer(-detector, names_to = "metric", values_to = "delta")
ov$delta[!is.finite(ov$delta)] <- NA
lim <- stats::quantile(abs(ov$delta), 0.95, na.rm = TRUE); if (!is.finite(lim) || lim == 0) lim <- 1
ovp <- ggplot(ov, aes(metric, detector, fill = pmax(pmin(delta, lim), -lim),
                      text = sprintf("%s | %s<br>%+.1f%% vs baseline", detector, metric, delta))) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_y_discrete(limits = rev(detector_channels)) +
  scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0,
                       name = "Δ% vs\nbaseline\n(red=worse)") +
  labs(x = NULL, y = NULL, title = sprintf("Detector performance change — %s vs baseline", latest)) +
  theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                  axis.text.x = element_text(angle = 30, hjust = 1))
gp(ovp, 680)
```

# Detail {.tabset}

## Peak clustering

The 6 URCP-38-2K populations are assigned **once per run** by model-based clustering
(flowClust, K = 6) on the asinh-transformed anchor detectors, then read out in every
detector. This is far more robust than per-detector peak finding: weak channels
can't resolve dim peaks alone, but the strong anchor detectors can, and a bead's
population label is fixed across all detectors. Labels are ordered 1–6 by ascending
brightness: peak 1 = blank particles, peaks 2–6 = fluorescent rainbow levels.
Clustering runs in `Process6PeakUltra.R`; this page reads the persisted result.

### 2D cluster assignment (V8-A × B9-A)

The per-run diagnostic: each run's gated beads shown on two detectors that both
resolve all 6 peaks. A clean assignment shows **6 tight, ordered, single-coloured
blobs** along the diagonal — blank (peak 1) at the dim end, brightest (peak 6) at
the top — with no interleaving.

```{r}
#| label: cluster2d-images
#| output: asis
clus_imgs <- sort(list.files("images/6peak_ultra_evo", pattern = "^cluster2d_.*[.]png$", full.names = TRUE))
if (length(clus_imgs) == 0) cat("_No saved cluster images yet — run `scripts/Process6PeakUltra.R`._\n\n") else {
  cat("::: {layout-ncol=3}\n")
  for (f in clus_imgs) cat(sprintf("![Run %s](../%s)\n\n", sub("^cluster2d_(.*)[.]png$","\\1",basename(f)), f))
  cat(":::\n")
}
```

### Cluster-size sanity

URCP-38-2K mixes the 6 populations in roughly equal proportions (blank particles
may be slightly more abundant). Cluster sizes should be broadly even — a quick
numeric check that no population was split or merged by the clustering step.

```{r}
#| label: cluster-sizes
#| fig-height: 3
csz <- pm |> group_by(run_label, peak) |>
  summarise(events = mean(n, na.rm = TRUE), .groups = "drop") |>
  mutate(peak = factor(peak, levels = 1:n_peaks))
gp(ggplot(csz, aes(factor(peak), events, fill = run_label,
                   text = sprintf("peak %s | %s<br>mean events %.0f", peak, run_label, events))) +
     geom_col(position = position_dodge()) +
     labs(title = "Cluster size per peak (events per run; peak 1 = blank)",
          x = "Peak", y = "Events", fill = NULL) +
     theme_bw(base_size = 10), 300)
```

## Tracking over time

Each peak's median intensity and %CV against acquisition timestamp — the QC payoff.
Stable lines mean stable performance; a step or drift flags a detector/laser
problem. Resolved, in-range peaks only. These fill in as runs accumulate (sparse,
mostly points, with few timestamps so far).

### Per-laser overview

The 64 detectors collapsed to a 5-laser × run heatmap of peak-6 deviation from each
detector's baseline (first run), averaged across the laser. A whole-row shift points
to a laser-level issue (excitation power, alignment) rather than one detector.

```{r}
#| label: laser-overview
#| fig-height: 3
p6_base <- pm |> filter(peak == n_peaks, in_range) |> group_by(detector) |>
  slice_min(run_label, n = 1, with_ties = FALSE) |>
  transmute(detector, base_median = median) |> ungroup()
laser_delta <- pm |> filter(peak == n_peaks, in_range) |>
  left_join(p6_base, by = "detector") |>
  mutate(delta_pct = (median - base_median) / base_median * 100,
         laser = factor(laser_of(detector), levels = laser_levels)) |>
  group_by(laser, run_label) |> summarise(delta_pct = mean(delta_pct, na.rm = TRUE), .groups = "drop")
lim <- max(abs(laser_delta$delta_pct), na.rm = TRUE); if (!is.finite(lim) || lim == 0) lim <- 1
gp(ggplot(laser_delta, aes(run_label, laser, fill = delta_pct,
        text = sprintf("%s laser | %s<br>%+.1f%% vs baseline", laser, run_label, delta_pct))) +
     geom_tile(colour = "white") +
     scale_y_discrete(limits = rev(laser_levels)) +
     scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0,
                          limits = c(-lim, lim), name = "Peak 6\nΔ vs t0") +
     labs(title = "Per-laser peak-6 deviation from baseline", x = NULL, y = "Laser") +
     theme_bw(base_size = 11) +
     theme(axis.text = element_text(size = 10)), 420)
```

::: {.callout-note collapse="false"}
### Per-peak median intensity by run

```{r}
#| label: trend-median
#| fig-height: 30
trend <- pm |> filter(resolved, in_range) |> mutate(peak = factor(peak, levels = 1:n_peaks))
ggplot(trend, aes(run_label, median, colour = peak, group = peak)) +
  geom_line() + geom_point(size = 0.8) +
  scale_y_flowjo_biexp(maxValue = biex_max, pos = biex_pos, neg = biex_neg, widthBasis = biex_width_basis) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(title = "Per-peak median intensity over time (peaks 1–6; 1 = blank)",
       x = NULL, y = "Median (biexp)", colour = "Peak") +
  theme_bw(base_size = 8) +
  theme(strip.text = element_text(size = 7), axis.text.x = element_text(size = 5, angle = 45, hjust = 1))
```
:::

::: {.callout-note collapse="false"}
### %robust-CV trend

```{r}
#| label: trend-cv
#| fig-height: 30
ggplot(trend, aes(run_label, rcv, colour = peak, group = peak)) +
  geom_line() + geom_point(size = 0.8) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  cv_y_scale +
  labs(title = "Per-peak robust %%CV over time (peaks 1–6; 1 = blank)",
       x = NULL, y = "Robust %%CV", colour = "Peak") +
  theme_bw(base_size = 8) +
  theme(strip.text = element_text(size = 7), axis.text.x = element_text(size = 5, angle = 45, hjust = 1))
```
:::

## Resolution & SI

Number of the 6 populations resolved per detector (0–6) and the staining index
(lowest resolved peak vs the blank). Detectors losing dim peaks over time = rising
noise floor.

```{r}
#| label: res-heat
gp(heat(res, "n_resolved", "Peaks resolved (0–6)"))
```

::: {layout-ncol=2}
```{r}
#| label: si-heat
gp(heat(si |> filter(is.finite(si)), "si", "Staining index", "absolute; higher = better"))
```

```{r}
#| label: si-heat-fc
gp(heat_fc(si |> filter(is.finite(si)), "si", "Staining index — fold-change", "vs its own best; red = degraded", higher_better = TRUE))
```
:::

## Staining index

Sensitivity over time: how far the **first resolved peak above the blank** sits
above peak 1 (blank particles), relative to the blank's spread — the classical
detector-sensitivity benchmark,
$\mathrm{SI} = (\widetilde{m}_{\text{pos}} - \widetilde{m}_1)/(2\,\widetilde{\sigma}_1)$.
Point colour marks which peak was used as the positive (it falls back from peak 2 toward
peak 6 in weak detectors); a change in that reference for a detector is worth a look.
A trend that drops = rising background or falling gain. The cross-detector snapshot
is the SI heatmap under **Resolution & SI**.

::: {.callout-note collapse="true"}
### SI trend per detector (expand — one panel per detector)

```{r}
#| label: si-trend
#| fig-height: 30
ggplot(si |> filter(!is.na(si)), aes(run_label, si)) +
  geom_line(aes(group = detector), colour = "grey55") +
  geom_point(aes(colour = factor(pos_peak)), size = 1.2) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(title = "Staining index over time (lowest resolved peak vs blank, peak 1)",
       x = NULL, y = "SI", colour = "Pos peak") +
  theme_bw(base_size = 8) +
  theme(strip.text = element_text(size = 7), axis.text.x = element_text(size = 5, angle = 45, hjust = 1))
```
:::

## %CV & noise floor

Peak 6 (brightest) uses **robust %CV** — it has a well-defined mean (~1–2% when
healthy). The blank (peak 1) uses its **absolute robust spread** (IQR/1.349), not
%CV: the blank sits near zero, so %CV explodes when the mean dips toward zero even
though the actual spread is unchanged. A rising blank spread is a genuine
noise-floor increase.

```{r}
#| label: cvtop-heat
gp(heat(cvtop, "cvt", "Peak-6 %CV", "brightest population (mean well-defined)"))
```

::: {layout-ncol=2}
```{r}
#| label: cv1-heat
gp(heat(cvp1, "sd1", "Peak-1 noise floor", "absolute robust spread (IQR/1.349), not %%CV"))
```

```{r}
#| label: cv1-heat-fc
gp(heat_fc(cvp1, "sd1", "Noise floor — fold-change", "lower is better; red = noisier than its best", higher_better = FALSE))
```
:::

```{r}
#| label: cvtop-trend
#| fig-height: 3.5
cl_lat <- cvtop |> group_by(laser = factor(laser_of(detector), levels = laser_levels), run_label) |>
  summarise(cvt = mean(cvt, na.rm = TRUE), .groups = "drop")
gp(ggplot(cl_lat, aes(run_label, cvt, colour = laser, group = laser)) +
     geom_line() + geom_point() +
     labs(title = "Mean peak-6 %%CV per laser (± control context as runs accumulate)",
          x = NULL, y = "Peak-6 %%CV") + theme_bw(base_size = 10) +
     theme(axis.text.x = element_text(angle = 30, hjust = 1)), 320)
```

## Brightest peak (peak 6) %CV

A focused per-detector view of peak 6 (the brightest fluorescent level) alone over
time. The brightest population is normally the tightest (~1–2%), so any rise is a
sensitive indicator of saturation, noise, or linearity issues at the top of the
dynamic range. The cross-detector snapshot and per-laser mean are under
**%CV & noise floor**.

::: {.callout-note collapse="true"}
### Peak-6 %CV trend per detector (expand — one panel per detector)

```{r}
#| label: peak6-cv-trend
#| fig-height: 30
peak6 <- pm |> filter(peak == n_peaks, in_range)
ggplot(peak6, aes(run_label, rcv, group = detector)) +
  geom_line(colour = "grey60") + geom_point(size = 1, colour = "firebrick") +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  cv_y_scale +
  labs(title = "Peak-6 robust %%CV over time (brightest URCP-38-2K level)",
       x = NULL, y = "Peak-6 %%CV") +
  theme_bw(base_size = 8) +
  theme(strip.text = element_text(size = 7), axis.text.x = element_text(size = 5, angle = 45, hjust = 1))
```
:::

## Spectral signature

Cosine similarity of each peak's normalised 64-detector signature to the baseline
run. 1.000 = unchanged; a drop indicates an optical / alignment / filter change
that per-detector intensities alone would miss.

```{r}
#| label: spec-heat
sp <- spec |> mutate(peak = factor(peak, levels = 1:n_peaks))
gp(ggplot(sp, aes(run_label, peak, fill = cosine,
                  text = sprintf("peak %s | %s<br>cosine %.4f", peak, run_label, cosine))) +
     geom_tile(colour = "white") +
     scale_fill_viridis_c(option = "magma", limits = c(min(0.99, min(sp$cosine, na.rm=TRUE)), 1)) +
     labs(title = "Spectral-signature similarity to baseline", x = NULL, y = "Peak") +
     theme_bw(base_size = 10) + theme(axis.text.x = element_text(angle = 30, hjust = 1)), 300)
```

## Per-run diagnostics

The bead gate saved by `Process6PeakUltra.R` per run — it should land on the dense
scatter cluster. (The 2D cluster-assignment diagnostic is in the **Peak clustering** tab.)

### Bead gate (FSC-A × SSC-A)

```{r}
#| label: gate-images
#| output: asis
gate_imgs <- sort(list.files("images/6peak_ultra_evo", pattern = "^gate_.*[.]png$", full.names = TRUE))
if (length(gate_imgs) == 0) cat("_No saved gate images yet — run `scripts/Process6PeakUltra.R`._\n\n") else {
  cat("::: {layout-ncol=3}\n")
  for (f in gate_imgs) cat(sprintf("![Run %s](../%s)\n\n", sub("^gate_(.*)[.]png$","\\1",basename(f)), f))
  cat(":::\n")
}
```

## Sensitivity (Q & B) · in development

::: {.callout-warning}
**In development.** The Q/B sensitivity model, MEF calibration, linearity, and gain
context below are still being validated against the URCP-38-2K bead lot — treat as
preliminary, not yet for routine QC decisions.
:::

**Q** = detection efficiency (photoelectrons per unit signal, gain-normalised) — a
drop means the detector collects less light per event. **B** = background/noise
term; **intrinsic CV** = the high-signal noise floor (bead + optics). Estimated
per detector from CV² vs 1/MFI across the resolved peaks. Because `Q = 1/slope`
blows up when the slope nears zero, fits that are **ill-conditioned** — near-flat
or negative slope, R² < 0.8, a non-monotonic CV profile, or a brightest-peak CV
> 15% — are **suppressed** (left blank) rather than reported as a phantom Q spike;
the underlying cause then shows in the %CV panel.

::: {.callout-note collapse="true"}
## Method & provenance — read me

This is the **Q–B detector-sensitivity model** of Hoffman & Wood (Wood JC,
*Cytometry* 1998;33:260–266; Wood & Hoffman, *Cytometry* 1998;33:256–259; Chase &
Hoffman, *Cytometry* 1998;33:267–279; Hoffman & Wood, *Current Protocols in
Cytometry* Unit 1.20, 2007), the same model behind Spherotech's **STN-17** and
BD's **CS&T Qr/Br**. Total spread decomposes as
`CV² = CV_intr² + (1/Q)(1/S) + (B/Q)(1/S²)`; fitting CV² vs 1/S gives intrinsic CV
(intercept), `1/Q` (linear term) and `B/Q` (curvature).

**The panel-wide values use gain-normalised MFI (`S = MFI/gain`), not MEF, so Q
and B are a *relative, non-normalised* index** — valid for **single-detector
trend monitoring (Levey-Jennings)**, *not* cross-detector comparison, and expect
small steps when a bead lot changes. Absolute, gain-independent sensitivity lives
in the **MEF-unit Q/B** and **detection-limit (MEF)** panels below, on the anchor
detectors.
:::

```{r}
#| label: fc-mode-banner
#| output: asis
if (fc_locked) cat(sprintf("> **FC reference: LOCKED** to qualified window %s – %s (per-detector median ± SD).\n",
                           fc_ref_window[1], fc_ref_window[2])) else
  cat(sprintf("> **FC reference: dynamic / preliminary** — each detector vs its own best of %d run(s). Lock a fixed reference once enough qualified runs accumulate.\n", n_runs))
```

::: {layout-ncol=2}
```{r}
#| label: q-heat
gp(heat(qb, "Q", "Detection efficiency Q", "absolute — gain-normalised; higher = better"))
```

```{r}
#| label: q-heat-fc
gp(heat_fc(qb, "Q", "Q — fold-change (per detector)", "vs its own best; red = degraded", higher_better = TRUE))
```
:::

**Background B** is shown absolute only: the relative B is a fit coefficient that can
be ≤ 0, so a fold-change vs best is undefined for it.

```{r}
#| label: b-heat
gp(heat(qb |> filter(is.finite(B)), "B", "Background B", "absolute — higher = more background/noise"))
```

```{r}
#| label: qb-flag-note
flagged <- qb |> filter(!is.na(qflag), qflag != "n<3")
if (nrow(flagged)) {
  fl <- flagged |> group_by(run_label) |>
    summarise(`detectors suppressed` = n(),
              `example reasons` = paste(head(sort(unique(qflag)), 3), collapse = " | "),
              .groups = "drop")
  knitr::kable(fl, caption = "Detector-runs where the Q/B fit was suppressed as ill-conditioned (blank tiles above) — NOT a sensitivity gain. The cause appears in the %%CV & noise-floor panel.")
}
```

```{r}
#| label: q-trend
#| fig-height: 3.5
ql <- qb |> mutate(laser = factor(laser_of(detector), levels = laser_levels)) |>
  group_by(laser, run_label) |> summarise(Q = median(Q, na.rm = TRUE), .groups = "drop")
gp(ggplot(ql, aes(run_label, Q, colour = laser, group = laser,
                  text = sprintf("%s laser | %s<br>median Q %.3g", laser, run_label, Q))) +
     geom_line() + geom_point() +
     labs(title = "Median detection efficiency per laser over time", x = NULL, y = "Q (median)") +
     theme_bw(base_size = 10) + theme(axis.text.x = element_text(angle = 30, hjust = 1)), 320)
```

True sensitivity in standardized units — the **detection limit (MEF)** per anchor
detector from the URCP-38-2K calibration. Gain-independent, so a rise is a real
sensitivity loss (not the QC gain loop).

```{r}
#| label: mef-lod-trend
#| eval: !expr has_mef
#| fig-height: 3.5
gp(ggplot(sens, aes(run_label, MEF_LOD, colour = detector, group = detector,
                    text = sprintf("%s (%s)<br>%s<br>MEF LOD %.0f", detector, fluor, run_label, MEF_LOD))) +
     geom_line() + geom_point() + scale_y_log10() +
     labs(title = "Detection limit (MEF) per anchor detector — URCP-38-2K",
          x = NULL, y = "MEF at detection threshold (log)") +
     theme_bw(base_size = 10) + theme(axis.text.x = element_text(angle = 30, hjust = 1)), 340)
```

### Q & B in MEF units (anchor detectors)

The Q/B heatmaps above are in **gain-normalised MFI** units — comparable across the
full 64-detector panel, but arbitrary. On the **anchor detectors** (where the
URCP-38-2K certificate gives a per-peak MEF) the *same* CV² regression is run
against true **MEF**, yielding Q and B in absolute, gain-independent units matching
the Spherotech **STN-17** definition. The **intrinsic CV** (high-signal floor) is
scale-invariant, so it should agree between the two fits — a built-in check.

```{r}
#| label: qb-mef-trend
#| eval: !expr has_mef
#| fig-height: 3.5
gp(ggplot(qb_mef, aes(run_label, Q, colour = detector, group = detector,
        text = sprintf("%s (%s)<br>%s<br>Q(MEF) %.3g | B(MEF) %.3g | intrinsic CV %.2f%%",
                       detector, fluor, run_label, Q, B, 100 * intr_cv))) +
     geom_line() + geom_point() + scale_y_log10() +
     labs(title = "Detection efficiency Q in MEF units — anchor detectors (STN-17)",
          x = NULL, y = "Q (MEF units, log)") +
     theme_bw(base_size = 10) + theme(axis.text.x = element_text(angle = 30, hjust = 1)), 340)
```

```{r}
#| label: qb-mef-table
#| eval: !expr has_mef
latest_mef <- max(qb_mef$run_label)
cmp <- qb_mef |> filter(run_label == latest_mef) |>
  transmute(Detector = detector, Fluor = fluor,
            `Q (MEF)` = signif(Q, 3), `B (MEF)` = signif(B, 3),
            `intr.CV %% (MEF fit)` = round(100 * intr_cv, 2)) |>
  left_join(qb |> mutate(detector = as.character(detector)) |> filter(run_label == latest_mef) |>
              transmute(Detector = detector,
                        `Q (relative)` = signif(Q, 3),
                        `intr.CV %% (rel. fit)` = round(100 * intr_cv, 2)),
            by = "Detector")
knitr::kable(cmp, caption = sprintf(
  "Anchor-detector sensitivity, run %s. The two intrinsic-CV columns should agree (scale-invariant); Q (MEF) is the absolute, gain-independent value, Q (relative) the gain-normalised one.", latest_mef))
```

The regression above uses every resolved peak. As an independent check, the
**literal STN-17 closed form** uses just three populations — the blank (peak 1),
the dimmest fluorescent bead carrying an assigned MEF (peak 3) and the brightest
(peak 6). The blank carries no assigned MEF, so it is **extrapolated through the
per-run calibration**, then `Q = 1/(CV²_blank × MEF_blank)`. The two Q estimates
won't match exactly (3-point algebra vs. multi-point fit) but should track each
other; large divergence flags a nonlinear detector.

```{r}
#| label: qb-stn17-table
#| eval: !expr has_mef
stn_tab <- stn17 |> filter(run_label == latest_mef) |>
  transmute(Detector = detector, Fluor = fluor,
            `MFI blank` = signif(m1, 3), `MEF blank (extrap.)` = signif(MEF_blank, 3),
            `CV%% blank` = round(100 * cv1, 1),
            `Q (STN-17)` = signif(Q_stn, 3), `B (STN-17)` = signif(B_stn, 3)) |>
  left_join(qb_mef |> filter(run_label == latest_mef) |>
              transmute(Detector = detector, `Q (MEF regr.)` = signif(Q, 3)),
            by = "Detector")
knitr::kable(stn_tab, caption = sprintf(
  "STN-17 closed form, run %s. MEF blank is extrapolated through the per-run calibration; Q (STN-17) is 1/(CV²_blank × MEF_blank), beside the multi-point regression Q.", latest_mef))
```

### Linearity

**MEF-calibrated** (SPHERO URCP-38-2K) for the 18 anchor detectors, plus a
rank-based R² overview for all 64. On the log–log fit a slope ≠ 1 or curvature
flags APD compression/saturation (bright end) or floor effects (dim end). The
certificate's 5 fluorescent peaks map directly to data peaks 2–6 (peak 1 = blank,
unassigned).

```{r}
#| label: lin-mef
#| eval: !expr has_mef
#| fig-height: 6.5
ld <- lin_mef |> filter(run_label == max(run_label)) |> mutate(panel = paste0(fluor, " (", detector, ")"))
gp(ggplot(ld, aes(MEF, median)) +
     geom_smooth(method = "lm", se = FALSE, colour = "grey65", linewidth = 0.4) +
     geom_point(aes(text = sprintf("data peak %s<br>MEF %.0f<br>MFI %.0f", peak, MEF, median)), size = 1.3) +
     scale_x_log10() + scale_y_log10() +
     facet_wrap(~ panel, scales = "free", ncol = 3) +
     labs(title = sprintf("MEF-calibrated linearity — %s", max(lin_mef$run_label)),
          x = "Assigned MEF (URCP-38-2K)", y = "Measured MFI") +
     theme_bw(base_size = 8), 600)
```

```{r}
#| label: lin-mef-table
#| eval: !expr has_mef
sens |> group_by(detector) |> arrange(run_label) |>
  summarise(Fluorophore = dplyr::first(fluor),
            `slope (~1)` = round(dplyr::last(b), 2),
            `R2` = round(dplyr::last(r2), 4),
            `MEF LOD` = round(dplyr::last(MEF_LOD)),
            `LOD d%% vs base` = round((dplyr::last(MEF_LOD) - dplyr::first(MEF_LOD)) / dplyr::first(MEF_LOD) * 100, 1),
            .groups = "drop") |>
  arrange(match(detector, detector_channels)) |>
  knitr::kable(caption = "Per-anchor linearity & detection limit (latest run). MEF LOD = smallest distinguishable equivalent-fluorophore amount (gain-independent); lower = more sensitive.")
```

#### Rank-based overview (all detectors)

```{r}
#| label: lin-heat
gp(heat(lin, "r2", "Linearity R² (log MFI vs peak rank)", "1 = perfectly log-linear"))
```

### Gain context

The daily-QC **gain** applied on each run (rising = APD/laser aging) and
**MFI/gain** (gain-removed bead intensity; should be flat if the bead + optics are
stable).

```{r}
#| label: gain-missing-note
#| output: asis
if (length(runs_no_gain)) cat(sprintf(
  "::: {.callout-warning appearance='simple'}\nNo daily-QC gain found in the archive for **%s** — these run(s) are shown blank on every gain-normalised panel (relative Q/B, MFI/gain, and the executive overview's Q & MFI rows) rather than mixed onto a raw-MFI scale. Their **gain-independent** metrics (MEF-unit Q/B, linearity, detection limit, resolution, %%CV, spectral signature) are unaffected and fully valid. Update `data/EVO/Archive/ArchivedDataEVO.csv` with the run's gain to populate the gain panels.\n:::\n\n",
  paste(runs_no_gain, collapse = ", ")))
```

```{r}
#| label: gain-heat
#| eval: !expr has_gain
gtab <- pm |> filter(peak == n_peaks) |> distinct(run_label, detector, gain)
gp(heat(gtab, "gain", "Applied gain (peak-6 row)", "from daily QC"))
```

::: {layout-ncol=2}
```{r}
#| label: mfinorm-heat
#| eval: !expr has_gain
gp(heat(ptop, "mfi_norm", "Peak-6 MFI / gain", "absolute — gain-removed intensity"))
```

```{r}
#| label: mfinorm-heat-fc
#| eval: !expr has_gain
gp(heat_fc(ptop, "mfi_norm", "MFI/gain — fold-change", "vs its own best; red = dimmer", higher_better = TRUE))
```
:::

```{r}
#| label: no-gain
#| eval: !expr '!has_gain'
#| output: asis
cat("_No matching daily-QC gain found for these runs (data/EVO/Archive/ArchivedDataEVO.csv) — gain context unavailable, so Q/B and MFI use raw MFI._\n")
```
