---
title: "8 Peak Bead Tracking — Aurora EVO (5L)"
subtitle: "Spectral QC: peak intensity tracking over time"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

Secondary QC for the Cytek Aurora EVO (5L): RCP-30-5A 8-peak rainbow beads are
acquired each run and tracked over time to catch instrument drift. The heavy
work — reading the raw FCS, gating the beads, clustering the 8 populations and
measuring each population in all 64 detectors — is done **once per run** by
`Process8Peak.R`, which appends the results to `peak_metrics.csv`. This page only
**reads that archive**, so it renders quickly and adds no load to the site build.
(For per-run gating/clustering diagnostics, render
`8peak_tracking_spectral/EVO_8 Peak Tracking.qmd` on demand.)

```{r}
#| label: load
library(ggcyto)   # scale_*_flowjo_biexp axis; pulls in flowCore, which masks dplyr verbs
library(ggplot2)
library(tidyr)
library(readr)
library(dplyr)    # load LAST so dplyr's filter()/select() win the namespace masking

# Display-only: read the archive produced by Process8Peak.R. No FCS, no flowClust.
results_file <- if (file.exists("8peak_tracking_spectral/peak_metrics.csv")) {
  "8peak_tracking_spectral/peak_metrics.csv"
} else "peak_metrics.csv"
combined <- read_csv(results_file, show_col_types = FALSE)
combined$date <- as.Date(combined$date)

# Display constants — must match Process8Peak.R.
n_peaks          <- 8
biex_max         <- 4194304
biex_pos         <- 5.62
biex_neg         <- 0
biex_width_basis <- -250

# Detector acquisition order (UV -> V -> B -> YG -> R), derived from the archive
# so this page needs no FCS to know the channel layout.
laser_levels <- c("UV", "V", "B", "YG", "R")
det_order <- combined |>
  distinct(detector) |>
  mutate(laser = sub("[0-9]+-A$", "", detector),
         num   = as.integer(sub("-A$", "", sub("^(UV|V|B|YG|R)", "", detector)))) |>
  arrange(match(laser, laser_levels), num)
detector_channels <- det_order$detector
combined$detector <- factor(combined$detector, levels = detector_channels)
```

## Run status

One-line health summary per run: how many of the 64 detectors had peak 8
**out of range** (saturated at the 4,194,304 ceiling). A non-zero count means
that detector's brightest signal can't be trusted on that day, and any
intensity/SI metric derived from it is suspect.

```{r}
#| label: run-status
#| results: asis

status <- combined |>
  filter(peak == n_peaks) |>
  group_by(date) |>
  summarise(saturated = sum(!in_range), total = n(), .groups = "drop") |>
  arrange(date)

for (i in seq_len(nrow(status))) {
  r    <- status[i, ]
  cls  <- if (r$saturated == 0) "callout-note" else "callout-warning"
  cat(sprintf("::: {.%s}\n**Run %s** — %d / %d detectors saturated at peak 8.\n:::\n\n",
              cls, format(r$date), r$saturated, r$total))
}
```

## Per-run diagnostics

These per-run images are saved by `Process8Peak.R` when each run is processed,
so the page stays light. The **bead gate** should land on the dense scatter
cluster (red, not the debris); the **2D cluster** view should show 8 tight,
ordered, single-coloured blobs along the diagonal.

### Bead gate (FSC-A × SSC-A)

```{r}
#| label: gate-images
#| output: asis
gate_imgs <- sort(list.files("images/8peak_evo", pattern = "^gate_.*[.]png$", full.names = TRUE))
if (length(gate_imgs) == 0) {
  cat("_No saved gate images yet — run `Process8Peak.R`._\n\n")
} else {
  cat("::: {layout-ncol=3}\n")
  for (f in gate_imgs) {
    d <- sub("^gate_(.*)[.]png$", "\\1", basename(f))
    cat(sprintf("![Run %s](../%s)\n\n", d, f))   # ../ : path is relative to the page in pages/
  }
  cat(":::\n")
}
```

### 2D cluster assignment (V1-A × YG1-A)

```{r}
#| label: cluster2d-images
#| output: asis
clus_imgs <- sort(list.files("images/8peak_evo", pattern = "^cluster2d_.*[.]png$", full.names = TRUE))
if (length(clus_imgs) == 0) {
  cat("_No saved cluster images yet — run `Process8Peak.R`._\n\n")
} else {
  cat("::: {layout-ncol=3}\n")
  for (f in clus_imgs) {
    d <- sub("^cluster2d_(.*)[.]png$", "\\1", basename(f))
    cat(sprintf("![Run %s](../%s)\n\n", d, f))   # ../ : path is relative to the page in pages/
  }
  cat(":::\n")
}
```

## Tracking over time

Each peak's median intensity and %CV plotted against acquisition date, per
detector, from the accumulated archive. Stable lines mean stable instrument
performance; a step or drift flags a detector/laser problem. Only resolved,
in-range peaks are shown. The two heatmaps give an executive overview before the
detailed per-detector trends.

### Per-laser overview

Collapses the 64 detectors into a 5-laser × dates heatmap of peak-8 deviation
from baseline (first run for each detector, averaged across that laser's
detectors). A whole-row colour shift indicates a laser-level problem rather than
a single detector. The first run is baseline so its row is 0% by construction.

```{r}
#| label: laser-heatmap
#| fig-height: 3.5

laser_of <- function(d) sub("[0-9]+-A$", "", as.character(d))

p8_base <- combined |>
  filter(peak == n_peaks, in_range) |>
  group_by(detector) |>
  slice_min(date, n = 1, with_ties = FALSE) |>
  select(detector, base_median = median) |>
  ungroup()

laser_delta <- combined |>
  filter(peak == n_peaks, in_range) |>
  left_join(p8_base, by = "detector") |>
  mutate(delta_pct = (median - base_median) / base_median * 100,
         laser    = factor(laser_of(detector),
                           levels = c("UV", "V", "B", "YG", "R"))) |>
  group_by(laser, date) |>
  summarise(delta_pct = mean(delta_pct, na.rm = TRUE),
            n_det     = n(), .groups = "drop")

lim <- max(abs(laser_delta$delta_pct), na.rm = TRUE)
ggplot(laser_delta, aes(x = factor(date), y = laser, fill = delta_pct)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = sprintf("%+.1f%%", delta_pct)), size = 3) +
  scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac",
                       midpoint = 0, limits = c(-lim, lim),
                       name = "Peak 8\nΔ vs t0") +
  labs(x = "Acquisition date", y = "Laser",
       title = "Per-laser peak-8 deviation from baseline",
       caption = "Cell = mean % change of peak 8 median from first run, across that laser's detectors (in-range only).") +
  theme_bw(base_size = 10)
```

### Resolution-count heatmap

For each detector × run, the number of peaks resolved (0–8). A row going darker
over time is the cleanest signal of a detector losing its dim peaks — i.e.
rising noise floor or falling gain.

```{r}
#| label: resolution-heatmap
#| fig-height: 14

resolution <- combined |>
  group_by(date, detector) |>
  summarise(n_resolved = sum(resolved), .groups = "drop")

ggplot(resolution, aes(x = factor(date), y = detector, fill = n_resolved)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_y_discrete(limits = rev(detector_channels)) +
  scale_fill_viridis_c(option = "viridis", limits = c(0, n_peaks),
                       name = "Peaks\nresolved") +
  labs(x = "Acquisition date", y = NULL,
       title = "Peaks resolved per detector per run",
       caption = "0 = no peaks separated above noise; 8 = full RCP-30-5A staircase resolved.") +
  theme_bw(base_size = 9) +
  theme(axis.text.y = element_text(size = 6))
```

```{r}
#| label: trends
#| fig-width: 12
#| fig-height: 32

# Trend of per-peak median intensity by acquisition date. Only resolved,
# in-range peaks are plotted (unresolved peaks carry no trustworthy signal).
trend <- combined |>
  filter(resolved, in_range) |>
  mutate(peak = factor(peak, levels = 1:n_peaks))

ggplot(trend, aes(x = date, y = median, colour = peak, group = peak)) +
  geom_line() +
  geom_point(size = 1) +
  scale_y_flowjo_biexp(maxValue = biex_max, pos = biex_pos,
                       neg = biex_neg, widthBasis = biex_width_basis) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(x = "Acquisition date", y = "Median intensity (biexponential scale)",
       colour = "Peak",
       title = "Per-peak median intensity over time",
       caption = "Resolved, in-range peaks only. Single run shown as points until more dates accumulate.") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 8),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```

```{r}
#| label: trends-cv
#| fig-width: 12
#| fig-height: 32

# %robust-CV trend: the tightness of each peak is a sensitive drift indicator.
ggplot(trend, aes(x = date, y = rcv, colour = peak, group = peak)) +
  geom_line() +
  geom_point(size = 1) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(x = "Acquisition date", y = "Robust %CV", colour = "Peak",
       title = "Per-peak %CV over time",
       caption = "Resolved, in-range peaks only.") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 8),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```

### Brightest peak (peak 8) %CV

The brightest population is normally the tightest (rCV ~1–2%), so any rise in its
%CV is a sensitive indicator of detector linearity, saturation, or noise at the
top of the dynamic range. One line per detector, in_range only.

```{r}
#| label: trends-peak8-cv
#| fig-width: 12
#| fig-height: 32

peak8 <- combined |>
  filter(peak == n_peaks, in_range)

ggplot(peak8, aes(x = date, y = rcv, group = detector)) +
  geom_line(colour = "grey60") +
  geom_point(size = 1.2, colour = "firebrick") +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(x = "Acquisition date", y = "Peak 8 robust %CV",
       title = "Brightest peak (peak 8) %CV over time",
       caption = "In-range peak 8 only.") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 8),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```

### Blank peak (peak 1) %CV

The blank/dimmest population's spread tracks the noise floor: a stable peak 1
%CV means consistent background; a rising one signals increased electronic
noise, optical scatter contamination, or compensation issues. One line per
detector, in_range only.

```{r}
#| label: trends-peak1-cv
#| fig-width: 12
#| fig-height: 32

peak1 <- combined |>
  filter(peak == 1, in_range)

ggplot(peak1, aes(x = date, y = rcv, group = detector)) +
  geom_line(colour = "grey60") +
  geom_point(size = 1.2, colour = "steelblue") +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(x = "Acquisition date", y = "Peak 1 robust %CV",
       title = "Blank peak (peak 1) %CV over time",
       caption = "In-range peak 1 only.") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 8),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```

## Staining index (peak 1 vs lowest resolved peak)

How far the **first resolved peak above the blank** sits above peak 1, relative
to peak 1's own spread. Peak 1 is the "negative"; the positive is the lowest
peak `> 1` that is resolved and in range (surfaced as `pos_peak` so any change is
visible).

$$\mathrm{SI} = \dfrac{\widetilde{m}_{\text{pos}} - \widetilde{m}_{1}}{2 \times \widetilde{\sigma}_{1}}$$

Medians ($\widetilde{m}$) and robust SD ($\widetilde{\sigma} = \mathrm{IQR}/1.349$) on raw intensity. A higher SI
means cleaner separation of the dimmest detectable signal from background; a
downward trend flags rising background or falling detector gain.

```{r}
#| label: si

si_per_detector <- function(g) {
  cand <- g$peak[g$peak > 1 & g$resolved & g$in_range]
  if (length(cand) == 0) {
    return(data.frame(pos_peak = NA_integer_,
                      median_pos = NA_real_,
                      median_1   = g$median[g$peak == 1],
                      rsd_1      = NA_real_, si = NA_real_))
  }
  p     <- min(cand)
  m1    <- g$median[g$peak == 1]
  rsd_1 <- abs(g$rcv[g$peak == 1] * m1 / 100)   # rSD recovered from rCV
  data.frame(pos_peak   = as.integer(p),
             median_pos = g$median[g$peak == p],
             median_1   = m1,
             rsd_1      = rsd_1,
             si         = (g$median[g$peak == p] - m1) / (2 * rsd_1))
}

si_summary <- combined |>
  group_by(file, date, detector) |>
  group_modify(~ si_per_detector(.x)) |>
  ungroup()

# Example: SI for a strong vs a weak detector across the available runs.
si_summary |>
  filter(detector %in% c("YG1-A", "B4-A", "V8-A", "UV1-A")) |>
  mutate(si = round(si, 2),
         median_pos = round(median_pos),
         median_1   = round(median_1),
         rsd_1      = round(rsd_1, 1)) |>
  arrange(detector, date)
```

```{r}
#| label: si-trend
#| fig-width: 12
#| fig-height: 32

# SI trend per detector. Point colour shows which "positive" peak the SI was
# computed against — runs where this changes for a given detector deserve
# closer inspection, since the metric becomes apples-to-oranges if the
# reference peak shifts.
ggplot(si_summary, aes(x = date, y = si)) +
  geom_line(aes(group = detector), colour = "grey50") +
  geom_point(aes(colour = factor(pos_peak)), size = 1.5) +
  facet_wrap(~ detector, scales = "free_y", ncol = 4) +
  labs(x = "Acquisition date", y = "Staining index (lowest resolved peak vs peak 1)",
       colour = "Pos peak",
       title = "Sensitivity over time (peak 1 vs lowest resolved peak)",
       caption = "SI = (median_pos - median_1) / (2 * rSD_1). NA panels: no peak above 1 is resolved & in range.") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 8),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```
