---
title: "Cytek Aurora EVO"
format:
  dashboard:
    orientation: columns
    scrolling: true
---
```{r}
#| include: false
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(ggplot2)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")

# Levey-Jennings control bands. QC_Plots() returns a list of trend ggplots
# (one per detector); this layers robust control limits *behind* the existing
# trace, QC-fail squares and engineer-visit verticals. Center/spread are robust
# (median +/- MAD-scaled SD); QC-fail points are excluded from the baseline but
# still plot. Limits are recomputed from the visible history at render time.
add_lj_bands <- function(p, k = c(2, 3), min_n = 8,
                         exclude_flagged = TRUE, clamp0 = FALSE) {
  ycol <- p$labels$title
  d <- p$data
  if (is.null(ycol) || !ycol %in% names(d) || !"DateTime" %in% names(d)) return(p)
  vals <- suppressWarnings(as.numeric(d[[ycol]]))
  keep <- is.finite(vals)
  fcol <- paste0("Flag-", ycol)
  if (exclude_flagged && fcol %in% names(d)) {
    fl <- suppressWarnings(as.logical(d[[fcol]]))
    keep <- keep & !(fl %in% TRUE)
  }
  base <- vals[keep]
  if (length(base) < min_n) return(p)                       # preliminary: no bands
  center <- stats::median(base, na.rm = TRUE)
  s <- stats::mad(base, constant = 1.4826, na.rm = TRUE)
  if (!is.finite(s) || s <= 0) return(p)                    # degenerate / flat
  k <- sort(unique(k)); inner <- k[1]; outer <- k[length(k)]
  xr <- range(d$DateTime, na.rm = TRUE)
  lo <- function(m) { v <- center - m * s; if (clamp0) max(v, 0) else v }
  hi <- function(m) center + m * s
  rect <- function(ymin, ymax, fillcol)
    geom_rect(data = data.frame(xmin = xr[1], xmax = xr[2], ymin = ymin, ymax = ymax),
              inherit.aes = FALSE,
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              fill = fillcol, alpha = 0.12)
  band_layers <- list(
    rect(lo(inner), hi(inner), "#2ca02c"),   # in-control (green)
    rect(lo(outer), lo(inner), "#ff7f0e"),   # warning below (amber)
    rect(hi(inner), hi(outer), "#ff7f0e"),   # warning above (amber)
    geom_hline(yintercept = center, color = "grey40", linewidth = 0.4),
    geom_hline(yintercept = c(lo(outer), hi(outer)),
               color = "#d62728", linetype = "dashed", linewidth = 0.3)
  )
  p$layers <- c(band_layers, p$layers)
  p
}

# Thin wrapper: add bands to every plot in a list, then make them interactive.
lj_render <- function(plotlist, clamp0 = FALSE)
  htmltools::tagList(lapply(lapply(plotlist, add_lj_bands, clamp0 = clamp0), ggplotly))

# The Aurora EVO runs SpectroFlo 4.x. Its Gain/rCV Archive (ArchivedDataEVO.csv)
# is produced from the DailyQC reports by the 4.x parser. The page renders once
# that archive is present, and shows an "awaiting data" placeholder until then.
ArchiveDir <- file.path(MainFolder, "EVO", "Archive")
GainFile <- list.files(ArchiveDir, pattern = "Archived", full.names = TRUE)
HasGain <- length(GainFile) == 1 && nrow(read.csv(GainFile[1], nrow = 1)) > 0
HasData <- HasGain
```

```{r}
#| eval: !expr '!HasData'
#| content: valuebox
#| title: "Cytek Aurora EVO"
#| icon: hourglass-split
#| color: secondary
list(value = "Awaiting first QC data")
```

```{r}
#| include: false
# No gain archive yet -> clean page (sidebar + placeholder above) and stop.
if (!HasData) knitr::knit_exit()
```

```{r}
Gain_EVO <- Luciernaga:::CurrentData(x="EVO", MainFolder=MainFolder, type = "Gain")
TheDate <- Gain_EVO %>% slice(1) %>% pull(DateTime) %>% as.Date()

WindowOfInterest <- Sys.time() - months(12)
Gain_EVO <- Gain_EVO %>% filter(DateTime >= WindowOfInterest)

Data <- read.csv("data/reference/AuroraMaintenance.csv", check.names=FALSE)
Data <- Data %>% filter(!str_detect(reason, "lean"))
RepairEVO <- Data %>% filter(instrument %in% "EVO")
```

```{r}
x <- Gain_EVO
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "Gain")]

UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)
LaserGains <- TheColumns[str_detect(TheColumns, "Laser")]
LaserDelayGains <- LaserGains[str_detect(LaserGains, "Delay")]
LaserDelayGains <- Luciernaga:::ColorPriority(LaserDelayGains)
LaserPowerGains <- LaserGains[str_detect(LaserGains, "Power")]
LaserPowerGains <- Luciernaga:::ColorPriority(LaserPowerGains)
ScalingGains <- TheColumns[str_detect(TheColumns, "Scaling")]
ScalingGains <- Luciernaga:::ColorPriority(ScalingGains)

UltraVioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

VioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict = TRUE, YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

BluePlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

YellowGreenPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

RedPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                     RepairVisits=RepairEVO)

ScatterPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)

LaserDelayPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserDelayGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)

LaserPowerPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserPowerGains,
                                plotType = "individual", returntype = "plots",
                                YAxisLabel = " ", RepairVisits=RepairEVO)

ScalingPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScalingGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)
```

```{r}
x <- Gain_EVO
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")

TheColumnNames <- TheColumns[str_detect(TheColumns, "rCV")]
UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

UltraVioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

VioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict=TRUE, YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

BluePlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

YellowGreenPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

RedPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                     RepairVisits=RepairEVO)

ScatterPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)
```

## {.sidebar}
Dashboard data for the **Cytek Aurora EVO** last updated on **`r TheDate`**

**Note:** The EVO runs SpectroFlo 4.x. Gain and %rCV are shown from the daily QC reports.

**First Column: Gain** Gain (Voltage) values set for instrument after QC. Changes over time reflective of laser health.
**Second Column: rCV** Percentage Robust Coefficient of Variation (rCV) after QC. Higher values reflect decreased resolution between positive and negative for that detector.

**Control bands:** each plot shows the robust centerline (median) with green (within ±2 SD) and amber (±2–3 SD) zones and red dashed action limits (±3 SD), computed from each detector's own history (robust SD = MAD; QC-fail points excluded from the baseline).

For additional information concerning individual parameter tabs, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek Aurora cytometers, adapted from the [UMGCC FCSS InstrumentQC](https://umgccfcss.github.io/InstrumentQC/) project.

This dashboard was created with [Quarto](https://quarto.org/), based on work by [David Rach](https://github.com/DavidRach).

## Gain {.tabset}

```{r}
#| title: UltraViolet
lj_render(UltraVioletPlotsGain)
```

```{r}
#| title: Violet
lj_render(VioletPlotsGain)
```

```{r}
#| title: Blue
lj_render(BluePlotsGain)
```

```{r}
#| title: YellowGreen
lj_render(YellowGreenPlotsGain)
```

```{r}
#| title: Red
lj_render(RedPlotsGain)
```

```{r}
#| title: Scatter
lj_render(ScatterPlotsGain)
```

```{r}
#| title: LaserPower
lj_render(LaserPowerPlotsGain)
```

```{r}
#| title: LaserDelay
lj_render(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render(ScalingPlotsGain)
```

## rCV {.tabset}

```{r}
#| title: UltraViolet
lj_render(UltraVioletPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Violet
lj_render(VioletPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Blue
lj_render(BluePlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: YellowGreen
lj_render(YellowGreenPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Red
lj_render(RedPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Scatter
lj_render(ScatterPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: LaserPower
lj_render(LaserPowerPlotsGain)
```

```{r}
#| title: LaserDelay
lj_render(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render(ScalingPlotsGain)
```
