---
title: "8 Peak Bead Tracking — Aurora EVO (5L)"
subtitle: "Spectral QC: instrument-performance tracking from RCP-30-5A beads"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    fig-width: 9
---

```{r}
#| label: load
#| include: false
library(ggcyto)   # scale_*_flowjo_biexp; pulls in flowCore (masks dplyr verbs)
library(ggplot2)
library(tidyr)
library(readr)
library(plotly)
library(dplyr)    # LAST so dplyr verbs win the namespace masking

results_file <- if (file.exists("8peak_tracking_spectral/peak_metrics.csv")) {
  "8peak_tracking_spectral/peak_metrics.csv"
} else "peak_metrics.csv"
pm <- read_csv(results_file, show_col_types = FALSE)
pm$date <- as.Date(pm$date)

n_peaks  <- 8
biex_max <- 4194304; biex_pos <- 5.62; biex_neg <- 0; biex_width_basis <- -250

# Detector acquisition order (UV -> V -> B -> YG -> R) from the archive.
laser_levels <- c("UV", "V", "B", "YG", "R")
laser_of <- function(d) sub("[0-9]+-A$", "", as.character(d))
det_order <- pm |> distinct(detector) |>
  mutate(laser = laser_of(detector),
         num   = as.integer(sub("-A$", "", sub("^(UV|V|B|YG|R)", "", detector)))) |>
  arrange(match(laser, laser_levels), num)
detector_channels <- det_order$detector
pm$detector <- factor(pm$detector, levels = detector_channels)
runs   <- sort(unique(pm$date)); n_runs <- length(runs); base_date <- runs[1]

# ---- GAIN cross-reference: the daily QC gain that was applied on each run ----
# On the Aurora, daily QC pins bead MFI to a target by adjusting APD gain, so raw
# MFI is largely held constant by design. Dividing MFI by gain de-pins it, giving
# a ~gain-independent signal; and CV-based sensitivity (Q/B) is gain-invariant.
gfile <- "data/EVO/Archive/ArchivedDataEVO.csv"
has_gain <- file.exists(gfile)
if (has_gain) {
  g <- read.csv(gfile, check.names = FALSE)
  g$Date <- as.Date(lubridate::ymd_hms(g$DateTime))
  glong <- g |>
    select(Date, ends_with("-Gain")) |>
    pivot_longer(-Date, names_to = "detg", values_to = "gain") |>
    mutate(detector = paste0(sub("-Gain$", "", detg), "-A")) |>
    group_by(Date, detector) |> summarise(gain = dplyr::first(gain), .groups = "drop")
  pm <- pm |> left_join(glong, by = c("date" = "Date", "detector"))
} else pm$gain <- NA_real_
pm <- pm |> mutate(mfi_norm = if_else(!is.na(gain) & gain > 0, median / gain, median))

# ---- per-detector x per-run metrics ----
# Sensitivity (Q & B): CV^2 = intrinsic_CV^2 + (1/Q)*(1/x) + (B/Q)*(1/x^2),
# where x is the per-event intensity. Two flavours of x:
#   - xvar = "mfi_norm" (default): gain-normalised MFI -> Q/B in arbitrary but
#     gain-stabilised units; available for all 64 detectors (relative drift).
#   - xvar = "MEF": Spherotech RCP-30-5A assigned MEF -> Q/B in true MEF units
#     (STN-17), gain-independent; anchor detectors only (absolute sensitivity).
# B reporting: the I(inv^2) coefficient is B/Q. true_b=TRUE divides by the linear
# coef (1/Q) to return B in physical x-units (STN-17-comparable); FALSE keeps the
# raw quadratic coef (the existing relative metric). intr_cv (the high-signal
# floor) is the intercept and is scale-invariant -> a cross-check between flavours.
# GUARD: Q = 1/slope explodes when the slope -> 0, so a noisy run can emit a phantom
# Q spike (e.g. 2026-05-29 R4-A read Q=35 under a red-laser fault). We suppress Q/B
# (-> NA, with a `qflag` reason) when the fit is ill-conditioned: near-flat/negative
# slope, poor R^2, a non-monotonic CV profile (CV^2 should rise with 1/x), or an
# inflated brightest-peak CV. The %CV & noise-floor panel still shows the real cause.
qb_fit <- function(d, xvar = "mfi_norm", true_b = FALSE,
                   min_r2 = 0.8, min_mono = 0.3, max_bright_cv = 15) {
  d <- d[d$resolved & d$in_range & d[[xvar]] > 0, ]
  if (nrow(d) < 3)
    return(tibble(intr_cv = NA_real_, Q = NA_real_, B = NA_real_, npts = nrow(d), qflag = "n<3"))
  cv2 <- (d$rcv / 100)^2; inv <- 1 / d[[xvar]]
  f2  <- lm(cv2 ~ inv); s1 <- coef(f2)[2]
  r2  <- summary(f2)$r.squared
  mono <- suppressWarnings(stats::cor(inv, cv2, method = "spearman"))   # CV^2 vs 1/x: want > 0
  bcv  <- d$rcv[which.max(d[[xvar]])]                                   # brightest in-range peak rCV
  intr <- sqrt(pmax(coef(f2)[1], 0))
  flags <- c(if (!is.finite(s1) || s1 <= 0)              "flat/neg slope"          else NULL,
             if (is.finite(r2)   && r2   < min_r2)        sprintf("R2=%.2f", r2)    else NULL,
             if (is.finite(mono) && mono < min_mono)      "non-monotonic"          else NULL,
             if (is.finite(bcv)  && bcv  > max_bright_cv) sprintf("brightCV=%.0f%%", bcv) else NULL)
  if (length(flags))
    return(tibble(intr_cv = intr, Q = NA_real_, B = NA_real_, npts = nrow(d),
                  qflag = paste(flags, collapse = "; ")))
  s2 <- if (nrow(d) >= 4) coef(lm(cv2 ~ inv + I(inv^2)))[3] else NA_real_
  B  <- if (true_b) s2 / s1 else s2
  tibble(intr_cv = intr, Q = 1 / s1, B = B, npts = nrow(d), qflag = NA_character_)
}
qb <- pm |> group_by(date, detector) |> group_modify(~ qb_fit(.x)) |> ungroup()

# Resolution, peak-8/peak-1 %CV, peak-8 gain-normalised intensity
res   <- pm |> group_by(date, detector) |> summarise(n_resolved = sum(resolved), .groups = "drop")
cvp8  <- pm |> filter(peak == n_peaks, in_range) |> select(date, detector, rcv) |> rename(cv8 = rcv)
cvp1  <- pm |> filter(peak == 1, in_range) |> mutate(sd1 = rcv / 100 * median) |>
  select(date, detector, sd1)   # blank NOISE FLOOR = absolute robust SD (IQR/1.349); %CV is unstable near a ~0 mean
p8    <- pm |> filter(peak == n_peaks, in_range) |> select(date, detector, median, gain, mfi_norm)

# Staining index (lowest resolved peak vs peak 1)
si_one <- function(g) {
  cand <- g$peak[g$peak > 1 & g$resolved & g$in_range]
  if (!length(cand)) return(tibble(si = NA_real_))
  p <- min(cand); m1 <- g$median[g$peak == 1]; rsd1 <- abs(g$rcv[g$peak == 1] * m1 / 100)
  tibble(si = (g$median[g$peak == p] - m1) / (2 * rsd1))
}
si <- pm |> group_by(date, detector) |> group_modify(~ si_one(.x)) |> ungroup()

# Linearity (rank-based until lot AS01 MESF values are supplied): measured MFI vs
# peak rank, per detector-run. Slope/R^2 flag compression/saturation.
lin <- pm |> filter(resolved, in_range) |> group_by(date, detector) |>
  filter(n() >= 3) |>
  summarise(r2 = summary(lm(log10(median) ~ peak))$r.squared, .groups = "drop")

# ---- MEF calibration (Spherotech RCP-30-5A lot AS01) for anchor detectors ----
mef_map <- tibble::tribble(
  ~fluor,                      ~detector,
  "Cascade Blue",              "V1-A",
  "Blue Fluorescent Protein",  "V3-A",
  "FITC",                      "B2-A",
  "PE",                        "YG1-A",
  "EDC",                       "YG3-A",
  "PE-Cy5",                    "YG5-A",
  "PE-Cy7",                    "YG9-A",
  "APC",                       "R1-A",
  "APC-Cy7",                   "R7-A")
mef_file <- "data/reference/RCP-30-5A_AS01_MEF.csv"
has_mef  <- file.exists(mef_file)
if (has_mef) {
  mefv <- readr::read_csv(mef_file, show_col_types = FALSE) |> inner_join(mef_map, by = "fluor")
  lin_mef <- pm |> mutate(detector = as.character(detector)) |>
    inner_join(mefv, by = c("detector", "peak")) |>
    filter(resolved, in_range, median > 0, MEF > 0)
  # log-log calibration per anchor x run (slope ~1 if linear; intercept holds the gain)
  calib <- lin_mef |> group_by(date, detector, fluor) |> filter(n() >= 3) |>
    summarise(b = coef(lm(log10(median) ~ log10(MEF)))[2],
              a = coef(lm(log10(median) ~ log10(MEF)))[1],
              r2 = summary(lm(log10(median) ~ log10(MEF)))$r.squared, .groups = "drop")
  # Detection limit in MEF: blank (peak 1) + 2 robust SD, inverted through the
  # calibration. The applied gain cancels, so MEF_LOD is gain-independent.
  blank <- pm |> filter(peak == 1) |> mutate(detector = as.character(detector)) |>
    semi_join(mef_map, by = "detector") |>
    transmute(date, detector, thr = median * (1 + 2 * rcv / 100))
  sens <- calib |> left_join(blank, by = c("date", "detector")) |>
    mutate(MEF_LOD = 10^((log10(thr) - a) / b))
  # MEF-unit Q & B (STN-17): same CV^2-vs-1/x regression as `qb`, but x = assigned
  # MEF instead of gain-normalised MFI, so Q/B are in true molecules-of-equivalent-
  # fluorophore units and fully gain-independent. Anchor detectors only (the ones
  # in mef_map). true_b returns B in MEF units; intr_cv should ~match the relative
  # fit's intr_cv on the same detector-run (scale-invariant cross-check).
  qb_mef <- lin_mef |> group_by(date, detector, fluor) |>
    group_modify(~ qb_fit(.x, xvar = "MEF", true_b = TRUE)) |> ungroup()
  # ---- Literal STN-17 closed form (independent cross-check of the regression) ----
  # The blank (peak 1) has no assigned MEF, so extrapolate it through the per-run
  # calibration: MEF_blank = 10^((log10(MFI_blank) - a) / b). Dim Bead 2 = peak 2
  # (assigned MEF); brightest = peak 8 (its CV^2 is the high-signal floor). Then
  #   Q = 1 / (CV^2_blank * MEF_blank)
  #   SD2 = MEF^2 * (CV^2 - CV^2_peak8);  B = (SD2_blank / SD2_dim2) * MEF_blank.
  p_cv <- function(p) pm |> filter(peak == p, median > 0) |>
    mutate(detector = as.character(detector)) |> semi_join(mef_map, by = "detector") |>
    transmute(date, detector, m = median, cv = rcv / 100)
  b1 <- p_cv(1)       |> rename(m1 = m, cv1 = cv)
  b2 <- p_cv(2)       |> rename(m2 = m, cv2 = cv) |>
    left_join(mefv |> filter(peak == 2) |> select(detector, MEF2 = MEF), by = "detector")
  b8 <- p_cv(n_peaks) |> transmute(date, detector, cv8 = cv)
  stn17 <- calib |> select(date, detector, fluor, a, b) |>
    left_join(b1, by = c("date", "detector")) |>
    left_join(b2, by = c("date", "detector")) |>
    left_join(b8, by = c("date", "detector")) |>
    mutate(
      MEF_blank = 10^((log10(m1) - a) / b),                  # blank MEF via calib
      Q_stn     = 1 / (cv1^2 * MEF_blank),
      SD2_blank = MEF_blank^2 * (cv1^2 - cv8^2),
      SD2_dim2  = MEF2^2      * (cv2^2 - cv8^2),
      B_stn     = (SD2_blank / SD2_dim2) * MEF_blank)
}

# Spectral signature drift: cosine similarity of the 64-detector signature of each
# peak vs the baseline run (gain-normalised). 1 = unchanged; drops = optical drift.
cos_sim <- function(a, b) sum(a * b) / (sqrt(sum(a * a)) * sqrt(sum(b * b)))
sig <- pm |> filter(in_range) |> select(date, peak, detector, mfi_norm) |>
  arrange(peak, detector)
spec <- sig |> group_by(peak) |> group_modify(function(g, k) {
  bv <- g |> filter(date == base_date)
  g |> group_by(date) |> summarise(
    cosine = cos_sim(mfi_norm, bv$mfi_norm[match(detector, bv$detector)]), .groups = "drop")
}) |> ungroup()

# Combined per-detector-run table + baseline deltas (for the overview heatmap)
metrics <- res |>
  left_join(qb,   by = c("date","detector")) |>
  left_join(cvp8, by = c("date","detector")) |>
  left_join(si,   by = c("date","detector")) |>
  left_join(p8 |> select(date, detector, mfi_norm), by = c("date","detector"))
add_base <- function(df, v) df |> group_by(detector) |> arrange(date) |>
  mutate("{v}_d" := (.data[[v]] - first(.data[[v]])) / abs(first(.data[[v]])) * 100) |> ungroup()
for (v in c("Q","B","cv8","si","mfi_norm","n_resolved")) metrics <- add_base(metrics, v)

# Control limits (mean +/- 2 SD) across runs, per detector, for the trend bands.
clim <- function(df, v) df |> group_by(detector) |>
  summarise(mu = mean(.data[[v]], na.rm = TRUE), sd = sd(.data[[v]], na.rm = TRUE), .groups = "drop") |>
  mutate(lcl = mu - 2*sd, ucl = mu + 2*sd)

# Helper: per-detector x date heatmap (rows = detector in acquisition order)
heat <- function(df, fill, title, sub = NULL, diverging = FALSE, limitsym = FALSE) {
  p <- ggplot(df, aes(factor(date), detector, fill = .data[[fill]],
                      text = sprintf("%s | %s<br>%s: %.3g", detector, date, fill, .data[[fill]]))) +
    geom_tile(colour = "white", linewidth = 0.1) +
    scale_y_discrete(limits = rev(detector_channels)) +
    labs(x = NULL, y = NULL, title = title, subtitle = sub, fill = "") +
    theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                    plot.subtitle = element_text(size = 8, colour = "grey40"))
  if (diverging) {
    lim <- if (limitsym) max(abs(df[[fill]]), na.rm = TRUE) else NA
    p <- p + scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0,
                                  limits = if (is.na(lim)) NULL else c(-lim, lim))
  } else p <- p + scale_fill_viridis_c(option = "viridis")
  p
}
gp <- function(p, h = 620) ggplotly(p, tooltip = "text", height = h) |>
  plotly::config(displayModeBar = FALSE)

# ---- Per-detector fold-change reference for the FC maps --------------------
# Absolute heatmaps span orders of magnitude across detectors, so a single linear
# colour scale is dominated by the brightest few. The companion "FC" maps show each
# detector relative to its OWN reference, removing absolute magnitude so dim and
# bright detectors are directly comparable. Two reference modes (fc_ref_window):
#   * NULL  -> DYNAMIC / preliminary: reference = each detector's own best run so far
#     (one-sided "fold vs best"). Fine while characterising, but it shifts as runs
#     accumulate (retroactively), so it can't anchor a stable trend.
#   * c("YYYY-MM-DD","YYYY-MM-DD") -> LOCKED: per-detector ROBUST CENTRE (median) + SD
#     over that qualification window (known-good runs only) are computed ONCE, frozen
#     to fc_ref_file, and the maps switch to a two-sided deviation-from-qualified-
#     normal view (red = worse, in log2 and SD units) so drift patterns / control
#     breaches emerge against a fixed anchor. Delete fc_ref_file to re-lock.
fc_ref_window <- NULL                                   # <- set a date range to LOCK
fc_ref_file   <- "data/reference/EVO_8peak_reference.csv"
fc_locked     <- !is.null(fc_ref_window)
fc_ref        <- NULL
if (fc_locked) {
  if (file.exists(fc_ref_file)) {
    fc_ref <- readr::read_csv(fc_ref_file, show_col_types = FALSE)
  } else {
    fc_metrics <- bind_rows(
      qb   |> transmute(date, detector, metric = "Q",        value = Q),
      si   |> transmute(date, detector, metric = "si",       value = si),
      cvp1 |> transmute(date, detector, metric = "sd1",      value = sd1),
      p8   |> transmute(date, detector, metric = "mfi_norm", value = mfi_norm))
    fc_ref <- fc_metrics |>
      filter(date >= as.Date(fc_ref_window[1]), date <= as.Date(fc_ref_window[2]), is.finite(value)) |>
      group_by(detector, metric) |>
      summarise(ref = median(value), ref_sd = sd(value), n_ref = dplyr::n(), .groups = "drop")
    readr::write_csv(fc_ref, fc_ref_file)
  }
}

# DYNAMIC reference: each detector vs its own best (one-sided). deg >= 0 (0 = best).
best_fc <- function(df, v, higher_better = TRUE) {
  df |> group_by(detector) |> mutate(
    .best = { x <- .data[[v]][is.finite(.data[[v]])]
              if (!length(x)) NA_real_ else if (higher_better) max(x) else min(x) },
    deg   = if (higher_better) log2(.best / .data[[v]]) else log2(.data[[v]] / .best),
    fold  = 2^deg) |> ungroup()
}
heat_fc <- function(df, v, title, sub = NULL, higher_better = TRUE) {
  if (fc_locked) {
    # LOCKED: two-sided deviation from the frozen qualified centre, oriented so +ve = worse
    r <- fc_ref |> filter(metric == v) |> transmute(detector = as.character(detector), ref, ref_sd)
    d <- df |> mutate(detector = as.character(detector)) |> left_join(r, by = "detector") |>
      mutate(value = .data[[v]],
             lfc   = log2(value / ref)        * (if (higher_better) -1 else 1),
             sddev = (value - ref) / ref_sd   * (if (higher_better) -1 else 1))
    hi <- stats::quantile(abs(d$lfc), 0.95, na.rm = TRUE); if (!is.finite(hi) || hi <= 0) hi <- 1
    ggplot(d, aes(factor(date), detector, fill = pmax(pmin(lfc, hi), -hi),
          text = sprintf("%s | %s<br>%s = %.3g<br>%+.2f log₂ vs normal | %+.1f SD",
                         detector, date, v, value, lfc, sddev))) +
      geom_tile(colour = "white", linewidth = 0.1) +
      scale_y_discrete(limits = rev(detector_channels)) +
      scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b", midpoint = 0,
                           limits = c(-hi, hi), name = "log₂ vs\nqualified\nnormal") +
      labs(x = NULL, y = NULL, title = title,
           subtitle = "deviation from locked reference (red = worse)") +
      theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                      plot.subtitle = element_text(size = 8, colour = "grey40"))
  } else {
    d  <- best_fc(df, v, higher_better)
    hi <- stats::quantile(d$deg, 0.95, na.rm = TRUE); if (!is.finite(hi) || hi <= 0) hi <- 1
    ggplot(d, aes(factor(date), detector, fill = pmin(deg, hi),
          text = sprintf("%s | %s<br>%s = %.3g<br>%.2f× its own best", detector, date, v, .data[[v]], fold))) +
      geom_tile(colour = "white", linewidth = 0.1) +
      scale_y_discrete(limits = rev(detector_channels)) +
      scale_fill_gradient(low = "white", high = "#b2182b", limits = c(0, hi),
                          name = "log₂ fold\nvs best", na.value = "grey92") +
      labs(x = NULL, y = NULL, title = title,
           subtitle = if (is.null(sub)) "preliminary: vs each detector's best so far" else sub) +
      theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                      plot.subtitle = element_text(size = 8, colour = "grey40"))
  }
}
```

Secondary QC for the Cytek Aurora EVO from RCP-30-5A 8-peak beads. **Note on the
Aurora:** daily QC adjusts each APD's *gain* to pin bead MFI to a target, so raw
MFI is held ~constant by design — the informative signals are the ones gain
can't compensate: **sensitivity (Q & B)**, **resolution**, **%CV**, **linearity**,
and **spectral-signature stability**. Heavy FCS processing is done once by
`scripts/Process8Peak.R`; this page reads the `peak_metrics.csv` archive (+ the
daily-QC gain) so it stays light.

```{r}
#| label: prelim-note
#| echo: false
#| output: asis
if (n_runs < 5) cat(sprintf(
  "::: {.callout-warning appearance='simple'}\nOnly **%d run(s)** so far — trends and the mean ± 2 SD control limits are **preliminary** and will firm up as more runs accumulate.\n:::\n\n", n_runs))
```

# Executive summary

```{r}
#| label: run-status
#| results: asis
status <- pm |> filter(peak == n_peaks) |> group_by(date) |>
  summarise(saturated = sum(!in_range), total = n(), .groups = "drop") |> arrange(date)
for (i in seq_len(nrow(status))) { r <- status[i, ]
  cat(sprintf("::: {.%s}\n**Run %s** — %d / %d detectors saturated at peak 8.\n:::\n\n",
              if (r$saturated == 0) "callout-note" else "callout-warning",
              format(r$date), r$saturated, r$total)) }
```

Latest run vs the baseline run (**`r format(base_date)`**), per detector. Blue = improved,
red = degraded (sign adjusted so red is always *worse*). Hover for values.

```{r}
#| label: overview-heat
latest <- max(runs)
ov <- metrics |> filter(date == latest) |>
  transmute(detector,
            `Sensitivity Q` = Q_d,                 # higher = better -> red when negative
            `Background B`  = -B_d,                 # higher B worse -> flip
            `Peak-8 %CV`    = -cv8_d,               # higher CV worse -> flip
            `Staining idx`  = si_d,
            `Peaks resolved`= n_resolved_d,
            `MFI/gain`      = mfi_norm_d) |>
  pivot_longer(-detector, names_to = "metric", values_to = "delta")
ov$delta[!is.finite(ov$delta)] <- NA
lim <- stats::quantile(abs(ov$delta), 0.95, na.rm = TRUE); if (!is.finite(lim) || lim == 0) lim <- 1
ovp <- ggplot(ov, aes(metric, detector, fill = pmax(pmin(delta, lim), -lim),
                      text = sprintf("%s | %s<br>%+.1f%% vs baseline", detector, metric, delta))) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_y_discrete(limits = rev(detector_channels)) +
  scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0,
                       name = "Δ% vs\nbaseline\n(red=worse)") +
  labs(x = NULL, y = NULL, title = sprintf("Detector performance change — %s vs baseline", format(latest))) +
  theme_bw(base_size = 8) + theme(axis.text.y = element_text(size = 5),
                                  axis.text.x = element_text(angle = 30, hjust = 1))
gp(ovp, 680)
```

# Detail {.tabset}

## Sensitivity (Q & B)

**Q** = detection efficiency (photoelectrons per unit signal, gain-normalised) — a
drop means the detector collects less light per event. **B** = background/noise
term; **intrinsic CV** = the high-signal noise floor (bead + optics). Estimated
per detector from CV² vs 1/MFI across the resolved peaks. Because `Q = 1/slope`
blows up when the slope nears zero, fits that are **ill-conditioned** — near-flat
or negative slope, R² < 0.8, a non-monotonic CV profile, or a brightest-peak CV
> 15% (e.g. a laser-instability run) — are **suppressed** (left blank) rather than
reported as a phantom Q spike; the underlying cause then shows in the %CV panel.

::: {.callout-note collapse="true"}
## Method & provenance — read me

This is the **Q–B detector-sensitivity model** of Hoffman & Wood (Wood JC,
*Cytometry* 1998;33:260–266; Wood & Hoffman, *Cytometry* 1998;33:256–259; Chase &
Hoffman, *Cytometry* 1998;33:267–279; Hoffman & Wood, *Current Protocols in
Cytometry* Unit 1.20, 2007), the same model behind Spherotech's **STN-17** and
BD's **CS&T Qr/Br** (BD app note *"Qr and Br in BD FACSDiva Software"*, 2012).
Total spread decomposes as `CV² = CV_intr² + (1/Q)(1/S) + (B/Q)(1/S²)`; fitting CV²
vs 1/S gives intrinsic CV (intercept), `1/Q` (linear term) and `B/Q` (curvature) —
algebraically identical to BD's `SD² = (1/Q)·S + B/Q` fit with `Q = 1/slope`,
`B = Q·intercept` (the MEF table's `true_b`), and to STN-17's closed form.

**These panel-wide values use gain-normalised MFI (`S = MFI/gain`), not MESF, so Q
and B are a *relative, non-normalised* index — not absolute MESF units.** This is
deliberate and is the intended use: BD reports its own Qr/Br in relative ABD units,
and explicitly states that *non-normalised* Qr "values are still valid and serve as
a powerful metric for tracking changes in detector performance over time" — a 50%
change in a non-normalised Qr equals a 50% change in a normalised one. So treat
these for **single-detector trend monitoring (Levey-Jennings)**, *not* for
cross-detector or cross-instrument comparison, and expect small steps when a bead
lot changes. Absolute, gain-independent sensitivity lives in the **MEF-unit Q/B**
and **detection-limit (MEF)** panels below, on the anchor detectors.
:::

**Fold-change maps:** beside each absolute map, a companion normalises every
detector to its own reference, stripping out absolute magnitude so dim and bright
detectors are directly comparable and the colour isn't dominated by the few
brightest. The reference has two modes (set by `fc_ref_window` in the load chunk):

- **Dynamic / preliminary (current):** reference = each detector's *best run so
  far*. Good while characterising, but it shifts retroactively as runs accumulate,
  so it can't anchor a stable trend.
- **Locked:** once ~8–10 qualified runs exist, set `fc_ref_window` to a date range
  of known-good runs. The per-detector **robust centre (median) ± SD** over that
  window is frozen to `data/reference/EVO_8peak_reference.csv`, and the maps switch
  to a two-sided **deviation-from-qualified-normal** view (red = worse, in log₂ and
  SD units) where real drift and control-limit breaches emerge against a fixed anchor.

```{r}
#| label: fc-mode-banner
#| output: asis
if (fc_locked) cat(sprintf("> **FC reference: LOCKED** to qualified window %s – %s (per-detector median ± SD). Maps show deviation from qualified normal.\n",
                           fc_ref_window[1], fc_ref_window[2])) else
  cat(sprintf("> **FC reference: dynamic / preliminary** — each detector vs its own best of %d run(s). Lock a fixed reference once enough qualified runs accumulate.\n", n_runs))
```

::: {layout-ncol=2}
```{r}
#| label: q-heat
gp(heat(qb, "Q", "Detection efficiency Q", "absolute — gain-normalised; higher = better"))
```

```{r}
#| label: q-heat-fc
gp(heat_fc(qb, "Q", "Q — fold-change (per detector)", "vs its own best; red = degraded", higher_better = TRUE))
```
:::

**Background B** is shown absolute only: the relative B is a fit coefficient that can
be ≤ 0, so a fold-change vs best is undefined for it.

```{r}
#| label: b-heat
gp(heat(qb |> filter(is.finite(B)), "B", "Background B", "absolute — higher = more background/noise"))
```

```{r}
#| label: qb-flag-note
flagged <- qb |> filter(!is.na(qflag), qflag != "n<3")
if (nrow(flagged)) {
  fl <- flagged |> group_by(date) |>
    summarise(`detectors suppressed` = n(),
              `example reasons` = paste(head(sort(unique(qflag)), 3), collapse = " | "),
              .groups = "drop")
  knitr::kable(fl, caption = "Detector-runs where the Q/B fit was suppressed as ill-conditioned or too noisy (blank tiles above) — NOT a sensitivity gain. The cause appears in the %CV & noise-floor panel below (e.g. the 2026-05-29 red-laser broadening).")
}
```

```{r}
#| label: q-trend
#| fig-height: 3.5
ql <- qb |> mutate(laser = factor(laser_of(detector), levels = laser_levels)) |>
  group_by(laser, date) |> summarise(Q = median(Q, na.rm = TRUE), .groups = "drop")
gp(ggplot(ql, aes(date, Q, colour = laser, group = laser,
                  text = sprintf("%s laser | %s<br>median Q %.3g", laser, date, Q))) +
     geom_line() + geom_point() +
     labs(title = "Median detection efficiency per laser over time", x = NULL, y = "Q (median)") +
     theme_bw(base_size = 10), 320)
```

True sensitivity in standardized units — the **detection limit (MEF)** per anchor
detector from the AS01 calibration. Gain-independent, so a rise is a real
sensitivity loss (not the QC gain loop).

```{r}
#| label: mef-lod-trend
#| eval: !expr has_mef
#| fig-height: 3.5
gp(ggplot(sens, aes(date, MEF_LOD, colour = detector, group = detector,
                    text = sprintf("%s (%s)<br>%s<br>MEF LOD %.0f", detector, fluor, date, MEF_LOD))) +
     geom_line() + geom_point() + scale_y_log10() +
     labs(title = "Detection limit (MEF) per anchor detector — lot AS01",
          x = NULL, y = "MEF at detection threshold (log)") +
     theme_bw(base_size = 10), 340)
```

### Q & B in MEF units (anchor detectors)

The Q/B heatmaps above are in **gain-normalised MFI** units — comparable across the
full 64-detector panel, but arbitrary. On the **anchor detectors** (where lot AS01
gives a per-peak MEF) the *same* CV² regression can be run against true **MEF**,
yielding Q and B in absolute, gain-independent units that match the Spherotech
**STN-17** definition. Because daily QC re-gains each detector, this MEF-anchored
sensitivity is the more defensible cross-day metric: it re-anchors to the bead
lot's physical scale rather than trusting the recorded gain. The **intrinsic CV**
(high-signal floor) is scale-invariant, so it should agree between the two fits —
a built-in consistency check.

```{r}
#| label: qb-mef-trend
#| eval: !expr has_mef
#| fig-height: 3.5
gp(ggplot(qb_mef, aes(date, Q, colour = detector, group = detector,
        text = sprintf("%s (%s)<br>%s<br>Q(MEF) %.3g | B(MEF) %.3g | intrinsic CV %.2f%%",
                       detector, fluor, date, Q, B, 100 * intr_cv))) +
     geom_line() + geom_point() + scale_y_log10() +
     labs(title = "Detection efficiency Q in MEF units — anchor detectors (STN-17)",
          x = NULL, y = "Q (MEF units, log)") +
     theme_bw(base_size = 10), 340)
```

```{r}
#| label: qb-mef-table
#| eval: !expr has_mef
latest <- max(qb_mef$date)
cmp <- qb_mef |> filter(date == latest) |>
  transmute(Detector = detector, Fluor = fluor,
            `Q (MEF)` = signif(Q, 3), `B (MEF)` = signif(B, 3),
            `intr.CV % (MEF fit)` = round(100 * intr_cv, 2)) |>
  left_join(qb |> mutate(detector = as.character(detector)) |> filter(date == latest) |>
              transmute(Detector = detector,
                        `Q (relative)` = signif(Q, 3),
                        `intr.CV % (rel. fit)` = round(100 * intr_cv, 2)),
            by = "Detector")
knitr::kable(cmp, caption = sprintf(
  "Anchor-detector sensitivity, run %s. The two intrinsic-CV columns should agree (scale-invariant); Q (MEF) is the absolute, gain-independent value, Q (relative) the gain-normalised one.", latest))
```

The regression above uses peaks 2–8 (every resolved peak). As an independent
check, the **literal STN-17 closed form** uses just three populations — the blank
(peak 1), one dim bead (peak 2) and the brightest (peak 8). The blank carries no
assigned MEF, so it is **extrapolated through the per-run calibration**,
`MEF_blank = 10^((log10(MFI_blank) − a) / b)`, then `Q = 1/(CV²_blank × MEF_blank)`.
The two Q estimates won't match exactly (3-point algebra vs. multi-point fit) but
should track each other; large divergence flags a nonlinear detector.

```{r}
#| label: qb-stn17-table
#| eval: !expr has_mef
stn_tab <- stn17 |> filter(date == latest) |>
  transmute(Detector = detector, Fluor = fluor,
            `MFI blank` = signif(m1, 3), `MEF blank (extrap.)` = signif(MEF_blank, 3),
            `CV% blank` = round(100 * cv1, 1),
            `Q (STN-17)` = signif(Q_stn, 3), `B (STN-17)` = signif(B_stn, 3)) |>
  left_join(qb_mef |> filter(date == latest) |>
              transmute(Detector = detector, `Q (MEF regr.)` = signif(Q, 3)),
            by = "Detector")
knitr::kable(stn_tab, caption = sprintf(
  "STN-17 closed form, run %s. MEF blank is extrapolated through the per-run calibration; Q (STN-17) is 1/(CV²_blank × MEF_blank), shown beside the multi-point regression Q for comparison.", latest))
```

## Resolution & SI

Number of the 8 peaks resolved per detector (0–8) and the staining index (lowest
resolved peak vs blank). Detectors losing dim peaks over time = rising noise floor.

```{r}
#| label: res-heat
gp(heat(res, "n_resolved", "Peaks resolved (0–8)"))
```

::: {layout-ncol=2}
```{r}
#| label: si-heat
gp(heat(si |> filter(is.finite(si)), "si", "Staining index", "absolute; higher = better"))
```

```{r}
#| label: si-heat-fc
gp(heat_fc(si |> filter(is.finite(si)), "si", "Staining index — fold-change", "vs its own best; red = degraded", higher_better = TRUE))
```
:::

## %CV & noise floor

Peak 8 (brightest) uses **robust %CV** — it has a well-defined mean (~1–2% when
healthy). The blank (peak 1) uses its **absolute robust spread** (IQR/1.349), not
%CV: the blank sits near zero, so %CV explodes when the mean dips toward zero even
though the actual spread is unchanged (e.g. UV8 on 2026-05-29). A rising blank
spread is a genuine noise-floor increase.

```{r}
#| label: cv8-heat
gp(heat(cvp8, "cv8", "Peak-8 %CV", "brightest population (mean well-defined)"))
```

::: {layout-ncol=2}
```{r}
#| label: cv1-heat
gp(heat(cvp1, "sd1", "Peak-1 noise floor", "absolute robust spread (IQR/1.349), not %CV"))
```

```{r}
#| label: cv1-heat-fc
gp(heat_fc(cvp1, "sd1", "Noise floor — fold-change", "lower is better; red = noisier than its best", higher_better = FALSE))
```
:::

```{r}
#| label: cv8-trend
#| fig-height: 3.5
cl <- clim(cvp8, "cv8")
cl_lat <- cvp8 |> group_by(laser = factor(laser_of(detector), levels = laser_levels), date) |>
  summarise(cv8 = mean(cv8, na.rm = TRUE), .groups = "drop")
band <- cvp8 |> mutate(laser = factor(laser_of(detector), levels = laser_levels)) |>
  group_by(laser) |> summarise(mu = mean(cv8, na.rm=TRUE), sd = sd(cv8, na.rm=TRUE), .groups="drop")
gp(ggplot(cl_lat, aes(date, cv8, colour = laser, group = laser)) +
     geom_line() + geom_point() +
     labs(title = "Mean peak-8 %CV per laser (± control context as runs accumulate)",
          x = NULL, y = "Peak-8 %CV") + theme_bw(base_size = 10), 320)
```

## Linearity

**MEF-calibrated** (Spherotech RCP-30-5A lot **AS01**) for the 9 anchor detectors,
plus a rank-based R² overview for all 64. On the log–log fit a slope ≠ 1 or
curvature flags APD compression/saturation (bright end) or floor effects (dim end).

```{r}
#| label: lin-mef
#| eval: !expr has_mef
#| fig-height: 6.5
ld <- lin_mef |> filter(date == max(date)) |> mutate(panel = paste0(fluor, " (", detector, ")"))
gp(ggplot(ld, aes(MEF, median)) +
     geom_smooth(method = "lm", se = FALSE, colour = "grey65", linewidth = 0.4) +
     geom_point(aes(text = sprintf("peak %s<br>MEF %.0f<br>MFI %.0f", peak, MEF, median)), size = 1.3) +
     scale_x_log10() + scale_y_log10() +
     facet_wrap(~ panel, scales = "free", ncol = 3) +
     labs(title = sprintf("MEF-calibrated linearity — %s", format(max(lin_mef$date))),
          x = "Assigned MEF (lot AS01)", y = "Measured MFI") +
     theme_bw(base_size = 8), 540)
```

```{r}
#| label: lin-mef-table
#| eval: !expr has_mef
sens |> group_by(detector) |> arrange(date) |>
  summarise(Fluorophore = dplyr::first(fluor),
            `slope (~1)` = round(dplyr::last(b), 2),
            `R2` = round(dplyr::last(r2), 4),
            `MEF LOD` = round(dplyr::last(MEF_LOD)),
            `LOD d% vs base` = round((dplyr::last(MEF_LOD) - dplyr::first(MEF_LOD)) / dplyr::first(MEF_LOD) * 100, 1),
            .groups = "drop") |>
  arrange(match(detector, detector_channels)) |>
  knitr::kable(caption = "Per-anchor linearity & detection limit (latest run). MEF LOD = smallest distinguishable equivalent-fluorophore amount (gain-independent); lower = more sensitive.")
```

### Rank-based overview (all detectors)

```{r}
#| label: lin-heat
gp(heat(lin, "r2", "Linearity R² (log MFI vs peak rank)", "1 = perfectly log-linear"))
```

## Gain context

The daily-QC **gain** applied on each run (rising = APD/laser aging — the main QC
dashboard's signal) and **MFI/gain** (gain-removed bead intensity; should be flat
if the bead + optics are stable).

```{r}
#| label: gain-heat
#| eval: !expr has_gain
gtab <- pm |> filter(peak == n_peaks) |> distinct(date, detector, gain)
gp(heat(gtab, "gain", "Applied gain (peak-8 row)", "from daily QC"))
```

::: {layout-ncol=2}
```{r}
#| label: mfinorm-heat
#| eval: !expr has_gain
gp(heat(p8, "mfi_norm", "Peak-8 MFI / gain", "absolute — gain-removed intensity"))
```

```{r}
#| label: mfinorm-heat-fc
#| eval: !expr has_gain
gp(heat_fc(p8, "mfi_norm", "MFI/gain — fold-change", "vs its own best; red = dimmer", higher_better = TRUE))
```
:::

```{r}
#| label: no-gain
#| eval: !expr '!has_gain'
#| output: asis
cat("_Gain archive (data/EVO/Archive/ArchivedDataEVO.csv) not found — gain context unavailable._\n")
```

## Spectral signature

Cosine similarity of each peak's normalised 64-detector signature to the baseline
run. 1.000 = unchanged; a drop indicates an optical / alignment / filter change
that per-detector intensities alone would miss.

```{r}
#| label: spec-heat
sp <- spec |> mutate(peak = factor(peak, levels = 1:n_peaks))
gp(ggplot(sp, aes(factor(date), peak, fill = cosine,
                  text = sprintf("peak %s | %s<br>cosine %.4f", peak, date, cosine))) +
     geom_tile(colour = "white") +
     scale_fill_viridis_c(option = "magma", limits = c(min(0.99, min(sp$cosine, na.rm=TRUE)), 1)) +
     labs(title = "Spectral-signature similarity to baseline", x = NULL, y = "Peak") +
     theme_bw(base_size = 10), 300)
```

## Per-run diagnostics

Saved by `Process8Peak.R` per run. The bead gate should land on the dense
scatter cluster; the 2D cluster view should show 8 tight, ordered blobs.

### Bead gate (FSC-A × SSC-A)

```{r}
#| label: gate-images
#| output: asis
gate_imgs <- sort(list.files("images/8peak_evo", pattern = "^gate_.*[.]png$", full.names = TRUE))
if (length(gate_imgs) == 0) cat("_No saved gate images yet — run `scripts/Process8Peak.R`._\n\n") else {
  cat("::: {layout-ncol=3}\n")
  for (f in gate_imgs) cat(sprintf("![Run %s](../%s)\n\n", sub("^gate_(.*)[.]png$","\\1",basename(f)), f))
  cat(":::\n")
}
```

### 2D cluster assignment (V1-A × YG1-A)

```{r}
#| label: cluster2d-images
#| output: asis
clus_imgs <- sort(list.files("images/8peak_evo", pattern = "^cluster2d_.*[.]png$", full.names = TRUE))
if (length(clus_imgs) == 0) cat("_No saved cluster images yet — run `scripts/Process8Peak.R`._\n\n") else {
  cat("::: {layout-ncol=3}\n")
  for (f in clus_imgs) cat(sprintf("![Run %s](../%s)\n\n", sub("^cluster2d_(.*)[.]png$","\\1",basename(f)), f))
  cat(":::\n")
}
```
