---
format:
  dashboard:
    orientation: columns
    scrolling: true
---
```{r}
#| message: FALSE
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(ggplot2)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")
TheList <- c("3L")

# Levey-Jennings control bands + drift-aware Westgard helpers (add_lj_bands,
# lj_render, westgard_eval, add_westgard, lj_render_wg, westgard_summary),
# shared with the EVO / 5L pages. See pages/_lj_helpers.R.
source("pages/_lj_helpers.R")

# Updating Data
# Data processing (parsing new QC .fcs/.csv files into the Archive) is handled
# by TheScript_3L.R, which runs on the instrument computer. This dashboard only
# *displays* the already-processed Archive data, so the parsing calls below are
# left commented out to keep the render reproducible and side-effect free.
# walk(.x=TheList, MainFolder=MainFolder, .f=Luciernaga:::DailyQCParse)
# walk(.x=TheList, .f=Luciernaga:::QCBeadParse, MainFolder=MainFolder)
```

```{r}
Gain_3L <- Luciernaga:::CurrentData(x="3L", MainFolder=MainFolder, type = "Gain")
TheDate <- Gain_3L %>% slice(1) %>% pull(DateTime) %>% as.Date()
```

```{r}
WindowOfInterest <- Sys.time() - months(12)

Gain_3L <- Gain_3L %>% filter(DateTime >= WindowOfInterest)
```

```{r}
Data <- read.csv("data/reference/AuroraMaintenance.csv", check.names=FALSE)

Data <- Data %>% filter(!str_detect(reason, "lean"))

Repair3L <- Data %>% filter(instrument %in% "3L")
```

```{r}
x <- Gain_3L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "Gain")]

VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)
LaserGains <- TheColumns[str_detect(TheColumns, "Laser")]
LaserDelayGains <- LaserGains[str_detect(LaserGains, "Delay")]
LaserDelayGains <- Luciernaga:::ColorPriority(LaserDelayGains)
ScalingGains <- TheColumns[str_detect(TheColumns, "Scaling")]
ScalingGains <- Luciernaga:::ColorPriority(ScalingGains)

VioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair3L)

BluePlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair3L)

RedPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                     RepairVisits=Repair3L)

ScatterPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)

LaserDelayPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserDelayGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)

ScalingPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScalingGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)
```

```{r}
x <- Gain_3L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "rCV")]

VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

VioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair3L)

BluePlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair3L)

RedPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                     RepairVisits=Repair3L)

ScatterPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)
```

## {.sidebar}
Dashboard data for the **Cytek Aurora 3L** last updated on **`r TheDate`**

**First Column: Gain** Gain (Voltage) values set for instrument after QC. Changes over time reflective of laser health.
**Second Column: rCV** Percentage Robust Coefficient of Variation (rCV) after QC. Higher values reflect decreased resolution between positive and negative for that detector.

**Control bands:** each plot shows the robust centerline (median) with green (within ±2 SD) and amber (±2–3 SD) zones and red dashed action limits (±3 SD), computed from each detector's own history (robust SD = MAD; QC-fail points excluded from the baseline).

**Westgard flags (Gain only):** hollow circles mark rule violations — red = 1₃s (>3 SD) or 2₂s (2 consecutive >2 SD, a step); amber = 1₂s (>2 SD warning). These use a **rolling ~45-day local baseline** (with a practical-significance floor), so they flag breaks from how the instrument runs *now*, independent of long-term aging — different from the full-window bands. Hover a marker for the rule; see the **Violations** tab for the list. The drift-counting rules (10x, 4₁s) are intentionally omitted because gain legitimately drifts as detectors age.

For additional information concerning individual parameter tabs, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek cytometers at the Sydney Cytometry Core Facility

For additional flow cytometry resources, visit our [website](https://sydneycytometry.org.au/)

This dashboard was created with [Quarto](https://quarto.org/), based on work by [David Rach](https://github.com/DavidRach).

## Gain {.tabset}

```{r}
#| title: Violet
lj_render_wg(VioletPlotsGain)
```

```{r}
#| title: Blue
lj_render_wg(BluePlotsGain)
```

```{r}
#| title: Red
lj_render_wg(RedPlotsGain)
```

```{r}
#| title: Scatter
lj_render_wg(ScatterPlotsGain)
```

```{r}
#| title: LaserDelay
lj_render_wg(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render_wg(ScalingPlotsGain)
```

```{r}
#| title: "Violations"
WGtab <- westgard_summary(c(VioletPlotsGain, BluePlotsGain, RedPlotsGain,
  ScatterPlotsGain, LaserDelayPlotsGain, ScalingPlotsGain))
if (nrow(WGtab) == 0) {
  cat("No Westgard violations in the current window.")
} else {
  knitr::kable(WGtab, row.names = FALSE,
    caption = paste("Westgard violations on Gain (most recent first).",
      "Rolling ~45-day local baseline; 1_3s = >3 SD (reject), 1_2s = >2 SD (warn),",
      "2_2s = 2 consecutive >2 SD same side (reject).",
      "Detectors with <8 recent points are unscored."))
}
```

## rCV {.tabset}

```{r}
#| title: Violet
lj_render(VioletPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Blue
lj_render(BluePlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Red
lj_render(RedPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Scatter
lj_render(ScatterPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: LaserDelay
lj_render(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render(ScalingPlotsGain)
```
