---
title: "Cytek Aurora EVO"
format:
  dashboard:
    orientation: columns
    scrolling: true
---
```{r}
#| include: false
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(ggplot2)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")

# Levey-Jennings control bands. QC_Plots() returns a list of trend ggplots
# (one per detector); this layers robust control limits *behind* the existing
# trace, QC-fail squares and engineer-visit verticals. Center/spread are robust
# (median +/- MAD-scaled SD); QC-fail points are excluded from the baseline but
# still plot. Limits are recomputed from the visible history at render time.
add_lj_bands <- function(p, k = c(2, 3), min_n = 8,
                         exclude_flagged = TRUE, clamp0 = FALSE) {
  ycol <- p$labels$title
  d <- p$data
  if (is.null(ycol) || !ycol %in% names(d) || !"DateTime" %in% names(d)) return(p)
  vals <- suppressWarnings(as.numeric(d[[ycol]]))
  keep <- is.finite(vals)
  fcol <- paste0("Flag-", ycol)
  if (exclude_flagged && fcol %in% names(d)) {
    fl <- suppressWarnings(as.logical(d[[fcol]]))
    keep <- keep & !(fl %in% TRUE)
  }
  base <- vals[keep]
  if (length(base) < min_n) return(p)                       # preliminary: no bands
  center <- stats::median(base, na.rm = TRUE)
  s <- stats::mad(base, constant = 1.4826, na.rm = TRUE)
  if (!is.finite(s) || s <= 0) return(p)                    # degenerate / flat
  k <- sort(unique(k)); inner <- k[1]; outer <- k[length(k)]
  xr <- range(d$DateTime, na.rm = TRUE)
  lo <- function(m) { v <- center - m * s; if (clamp0) max(v, 0) else v }
  hi <- function(m) center + m * s
  rect <- function(ymin, ymax, fillcol)
    geom_rect(data = data.frame(xmin = xr[1], xmax = xr[2], ymin = ymin, ymax = ymax),
              inherit.aes = FALSE,
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              fill = fillcol, alpha = 0.12)
  band_layers <- list(
    rect(lo(inner), hi(inner), "#2ca02c"),   # in-control (green)
    rect(lo(outer), lo(inner), "#ff7f0e"),   # warning below (amber)
    rect(hi(inner), hi(outer), "#ff7f0e"),   # warning above (amber)
    geom_hline(yintercept = center, color = "grey40", linewidth = 0.4),
    geom_hline(yintercept = c(lo(outer), hi(outer)),
               color = "#d62728", linetype = "dashed", linewidth = 0.3)
  )
  p$layers <- c(band_layers, p$layers)
  p
}

# Thin wrapper: add bands to every plot in a list, then make them interactive.
lj_render <- function(plotlist, clamp0 = FALSE)
  htmltools::tagList(lapply(lapply(plotlist, add_lj_bands, clamp0 = clamp0), ggplotly))

# ---- Westgard-style rule flagging (Gain only) --------------------------------
# Reduced, drift-aware ruleset. EVO Gain is a CONTROLLED output (the daily-QC
# routine sets the voltage to park the bead peak at target), so it drifts up as
# detectors age -- that is correct, not a fault. So we DROP the drift-counting
# rules (10x, 4_1s) and R_4s (needs two control levels; we have one/day), and keep
# 1_3s (reject), 1_2s (warn) and 2_2s (sustained step). Each point is scored
# against a ROLLING ~45-day baseline BEFORE it (not the full-window band stats):
# z = (x - local median) / s_eff, where s_eff is floored at the detector's global
# robust SD AND a practical-significance fraction of the level (rel_floor), so a
# trivial gain nudge on a rock-stable detector isn't called "2 SD". Returns the
# plot data with z / loc_center / wg_rules / wg_sev / wg_label, or NULL.
westgard_eval <- function(p, trail_days = 45, min_n = 8, rel_floor = 0.025) {
  ycol <- p$labels$title
  d <- p$data
  if (is.null(ycol) || !ycol %in% names(d) || !"DateTime" %in% names(d)) return(NULL)
  d <- d[order(d$DateTime), , drop = FALSE]
  x  <- suppressWarnings(as.numeric(d[[ycol]]))
  dt <- as.POSIXct(d$DateTime)
  fcol <- paste0("Flag-", ycol)
  flagged <- if (fcol %in% names(d)) (suppressWarnings(as.logical(d[[fcol]])) %in% TRUE)
             else rep(FALSE, nrow(d))
  n <- nrow(d)
  z <- rep(NA_real_, n); loc <- rep(NA_real_, n)
  gmad <- stats::mad(x[!flagged & is.finite(x)], constant = 1.4826)   # characteristic noise
  if (is.finite(gmad) && gmad > 0) {
    win <- trail_days * 86400
    for (i in seq_len(n)) {
      if (!is.finite(x[i])) next
      sel <- which(dt >= (dt[i] - win) & dt < dt[i] & !flagged & is.finite(x))
      if (length(sel) < min_n) next
      m <- stats::median(x[sel]); s <- stats::mad(x[sel], constant = 1.4826)
      s_eff <- max(s, gmad, rel_floor * abs(m))
      loc[i] <- m; z[i] <- (x[i] - m) / s_eff
    }
  }
  rule <- rep(NA_character_, n); sev <- rep(NA_character_, n)
  for (i in seq_len(n)) {
    if (is.na(z[i])) next
    rs <- character(0)
    if (abs(z[i]) > 3) rs <- c(rs, "1_3s") else if (abs(z[i]) > 2) rs <- c(rs, "1_2s")
    prev <- tail(which(!is.na(z[seq_len(i - 1)])), 1)                  # last scored point
    if (length(prev) == 1 &&
        ((z[i] > 2 && z[prev] > 2) || (z[i] < -2 && z[prev] < -2)))
      rs <- c(rs, "2_2s")
    if (length(rs) > 0) {
      rule[i] <- paste(rs, collapse = "+")
      sev[i]  <- if (any(rs %in% c("1_3s", "2_2s"))) "reject" else "warn"
    }
  }
  d$z <- z; d$loc_center <- loc; d$wg_rules <- rule; d$wg_sev <- sev
  d$wg_label <- ifelse(is.na(rule), NA_character_,
    sprintf("%s | %s | %.1f | z=%.2f | %s", ycol,
            format(as.Date(dt), "%Y-%m-%d"), x, z, rule))
  d
}

# Overlay Westgard violation markers ON TOP of the LJ bands for one plot.
add_westgard <- function(p, clamp0 = FALSE) {
  p <- add_lj_bands(p, clamp0 = clamp0)
  e <- westgard_eval(p)
  if (is.null(e)) return(p)
  v <- e[!is.na(e$wg_rules), , drop = FALSE]
  if (nrow(v) == 0) return(p)
  ycol <- p$labels$title
  v$.sev <- factor(v$wg_sev, levels = c("warn", "reject"))
  p + geom_point(data = v, inherit.aes = FALSE,
        aes(x = DateTime, y = .data[[ycol]], color = .sev, text = wg_label),
        shape = 1, stroke = 1.1, size = 4) +
      scale_color_manual(values = c(warn = "#ff7f0e", reject = "#d62728"),
                         drop = FALSE, guide = "none")
}

# Render Gain plot list with bands + Westgard markers (interactive).
lj_render_wg <- function(plotlist)
  htmltools::tagList(lapply(lapply(plotlist, add_westgard), ggplotly,
                            tooltip = c("x", "y", "text")))

# Collect violations across a set of detector plots into one recent-first table.
westgard_summary <- function(plotlist) {
  rows <- lapply(plotlist, function(p) {
    e <- westgard_eval(p); if (is.null(e)) return(NULL)
    v <- e[!is.na(e$wg_rules), , drop = FALSE]
    if (nrow(v) == 0) return(NULL)
    data.frame(Detector = p$labels$title, Date = as.Date(v$DateTime),
               Rule = v$wg_rules, Value = round(suppressWarnings(as.numeric(v[[p$labels$title]])), 1),
               z = round(v$z, 2), Severity = v$wg_sev, stringsAsFactors = FALSE)
  })
  out <- do.call(rbind, rows)
  if (is.null(out)) return(data.frame(Detector = character(), Date = as.Date(character()),
                                      Rule = character(), Value = numeric(),
                                      z = numeric(), Severity = character()))
  out[order(out$Date, decreasing = TRUE), , drop = FALSE]
}

# The Aurora EVO runs SpectroFlo 4.x. Its Gain/rCV Archive (ArchivedDataEVO.csv)
# is produced from the DailyQC reports by the 4.x parser. The page renders once
# that archive is present, and shows an "awaiting data" placeholder until then.
ArchiveDir <- file.path(MainFolder, "EVO", "Archive")
GainFile <- list.files(ArchiveDir, pattern = "Archived", full.names = TRUE)
HasGain <- length(GainFile) == 1 && nrow(read.csv(GainFile[1], nrow = 1)) > 0
HasData <- HasGain
```

```{r}
#| eval: !expr '!HasData'
#| content: valuebox
#| title: "Cytek Aurora EVO"
#| icon: hourglass-split
#| color: secondary
list(value = "Awaiting first QC data")
```

```{r}
#| include: false
# No gain archive yet -> clean page (sidebar + placeholder above) and stop.
if (!HasData) knitr::knit_exit()
```

```{r}
Gain_EVO <- Luciernaga:::CurrentData(x="EVO", MainFolder=MainFolder, type = "Gain")
TheDate <- Gain_EVO %>% slice(1) %>% pull(DateTime) %>% as.Date()

WindowOfInterest <- Sys.time() - months(12)
Gain_EVO <- Gain_EVO %>% filter(DateTime >= WindowOfInterest)

Data <- read.csv("data/reference/AuroraMaintenance.csv", check.names=FALSE)
Data <- Data %>% filter(!str_detect(reason, "lean"))
RepairEVO <- Data %>% filter(instrument %in% "EVO")
```

```{r}
x <- Gain_EVO
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "Gain")]

UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)
LaserGains <- TheColumns[str_detect(TheColumns, "Laser")]
LaserDelayGains <- LaserGains[str_detect(LaserGains, "Delay")]
LaserDelayGains <- Luciernaga:::ColorPriority(LaserDelayGains)
LaserPowerGains <- LaserGains[str_detect(LaserGains, "Power")]
LaserPowerGains <- Luciernaga:::ColorPriority(LaserPowerGains)
ScalingGains <- TheColumns[str_detect(TheColumns, "Scaling")]
ScalingGains <- Luciernaga:::ColorPriority(ScalingGains)

UltraVioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

VioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict = TRUE, YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

BluePlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

YellowGreenPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=RepairEVO)

RedPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                     RepairVisits=RepairEVO)

ScatterPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)

LaserDelayPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserDelayGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)

LaserPowerPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserPowerGains,
                                plotType = "individual", returntype = "plots",
                                YAxisLabel = " ", RepairVisits=RepairEVO)

ScalingPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScalingGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)
```

```{r}
x <- Gain_EVO
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")

TheColumnNames <- TheColumns[str_detect(TheColumns, "rCV")]
UltraVioletGains <- TheColumnNames[str_detect(TheColumnNames, "^UV")]
VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
YellowGreenGains <- TheColumnNames[str_detect(TheColumnNames, "^YG")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

UltraVioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=UltraVioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

VioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", strict=TRUE, YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

BluePlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

YellowGreenPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=YellowGreenGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=RepairEVO)

RedPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                     RepairVisits=RepairEVO)

ScatterPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=RepairEVO)
```

## {.sidebar}
Dashboard data for the **Cytek Aurora EVO** last updated on **`r TheDate`**

**Note:** The EVO runs SpectroFlo 4.x. Gain and %rCV are shown from the daily QC reports.

**First Column: Gain** Gain (Voltage) values set for instrument after QC. Changes over time reflective of laser health.
**Second Column: rCV** Percentage Robust Coefficient of Variation (rCV) after QC. Higher values reflect decreased resolution between positive and negative for that detector.

**Control bands:** each plot shows the robust centerline (median) with green (within ±2 SD) and amber (±2–3 SD) zones and red dashed action limits (±3 SD), computed from each detector's own history (robust SD = MAD; QC-fail points excluded from the baseline).

**Westgard flags (Gain only):** hollow circles mark rule violations — red = 1₃s (>3 SD) or 2₂s (2 consecutive >2 SD, a step); amber = 1₂s (>2 SD warning). These use a **rolling ~45-day local baseline** (with a practical-significance floor), so they flag breaks from how the instrument runs *now*, independent of long-term aging — different from the full-window bands. Hover a marker for the rule; see the **Violations** tab for the list. The drift-counting rules (10x, 4₁s) are intentionally omitted because gain legitimately drifts as detectors age.

For additional information concerning individual parameter tabs, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek Aurora cytometers, adapted from the [UMGCC FCSS InstrumentQC](https://umgccfcss.github.io/InstrumentQC/) project.

This dashboard was created with [Quarto](https://quarto.org/), based on work by [David Rach](https://github.com/DavidRach).

## Gain {.tabset}

```{r}
#| title: UltraViolet
lj_render_wg(UltraVioletPlotsGain)
```

```{r}
#| title: Violet
lj_render_wg(VioletPlotsGain)
```

```{r}
#| title: Blue
lj_render_wg(BluePlotsGain)
```

```{r}
#| title: YellowGreen
lj_render_wg(YellowGreenPlotsGain)
```

```{r}
#| title: Red
lj_render_wg(RedPlotsGain)
```

```{r}
#| title: Scatter
lj_render_wg(ScatterPlotsGain)
```

```{r}
#| title: LaserPower
lj_render_wg(LaserPowerPlotsGain)
```

```{r}
#| title: LaserDelay
lj_render_wg(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render_wg(ScalingPlotsGain)
```

```{r}
#| title: "Violations"
WGtab <- westgard_summary(c(UltraVioletPlotsGain, VioletPlotsGain, BluePlotsGain,
  YellowGreenPlotsGain, RedPlotsGain, ScatterPlotsGain, LaserPowerPlotsGain,
  LaserDelayPlotsGain, ScalingPlotsGain))
if (nrow(WGtab) == 0) {
  cat("No Westgard violations in the current window.")
} else {
  knitr::kable(WGtab, row.names = FALSE,
    caption = paste("Westgard violations on Gain (most recent first).",
      "Rolling ~45-day local baseline; 1_3s = >3 SD (reject), 1_2s = >2 SD (warn),",
      "2_2s = 2 consecutive >2 SD same side (reject).",
      "Detectors with <8 recent points are unscored."))
}
```

## rCV {.tabset}

```{r}
#| title: UltraViolet
lj_render(UltraVioletPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Violet
lj_render(VioletPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Blue
lj_render(BluePlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: YellowGreen
lj_render(YellowGreenPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Red
lj_render(RedPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: Scatter
lj_render(ScatterPlotsRCV, clamp0 = TRUE)
```

```{r}
#| title: LaserPower
lj_render(LaserPowerPlotsGain)
```

```{r}
#| title: LaserDelay
lj_render(LaserDelayPlotsGain)
```

```{r}
#| title: LaserScaling
lj_render(ScalingPlotsGain)
```
