---
format:
  dashboard:
    orientation: columns
    scrolling: true
---
```{r}
#| message: FALSE
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")
TheList <- c("3L")

# Updating Data
# Data processing (parsing new QC .fcs/.csv files into the Archive) is handled
# by TheScript_3L.R, which runs on the instrument computer. This dashboard only
# *displays* the already-processed Archive data, so the parsing calls below are
# left commented out to keep the render reproducible and side-effect free.
# walk(.x=TheList, MainFolder=MainFolder, .f=Luciernaga:::DailyQCParse)
# walk(.x=TheList, .f=Luciernaga:::QCBeadParse, MainFolder=MainFolder)
```

```{r}
MFI_3L <- Luciernaga:::CurrentData(x="3L", MainFolder=MainFolder, type = "MFI")
Gain_3L <- Luciernaga:::CurrentData(x="3L", MainFolder=MainFolder, type = "Gain")
TheDate <- MFI_3L %>% slice(1) %>% pull(DATE)
```

```{r}
WindowOfInterest <- Sys.time() - months(12)

MFI_3L <- MFI_3L %>% filter(DateTime >= WindowOfInterest)
Gain_3L <- Gain_3L %>% filter(DateTime >= WindowOfInterest)
```

```{r}
Data <- read.csv("AuroraMaintenance.csv", check.names=FALSE)

Data <- Data %>% filter(!str_detect(reason, "lean"))

Repair3L <- Data %>% filter(instrument %in% "3L")
```

```{r}
x <- MFI_3L
x <- x %>% dplyr::filter(Timepoint %in% c("Before", "After"))
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheIntermediate <- TheColumns[!str_detect(TheColumns, "Gain")]
TheIntermediate <- TheIntermediate[!str_detect(TheIntermediate, "rCV")]
TheColumnNames <- TheIntermediate[str_detect(TheIntermediate, "-A")]

VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheIntermediate[str_detect(TheIntermediate, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

VioletPlotsMFI <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "comparison", returntype = "plots",
                      Metadata="Timepoint", strict = TRUE, YAxisLabel = "MFI",
                      RepairVisits=Repair3L)

BluePlotsMFI <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "comparison", returntype = "plots",
                      Metadata="Timepoint", strict = TRUE, YAxisLabel = "MFI",
                      RepairVisits=Repair3L)

RedPlotsMFI <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "comparison", returntype = "plots",
                     Metadata="Timepoint", strict = TRUE, YAxisLabel = "MFI",
                     RepairVisits=Repair3L)

ScatterPlotsMFI <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "comparison", returntype = "plots",
                     Metadata="Timepoint", strict = TRUE, YAxisLabel = " ",
                     RepairVisits=Repair3L)
```

```{r}
x <- Gain_3L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "Gain")]

VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)
LaserGains <- TheColumns[str_detect(TheColumns, "Laser")]
LaserDelayGains <- LaserGains[str_detect(LaserGains, "Delay")]
LaserDelayGains <- Luciernaga:::ColorPriority(LaserDelayGains)
ScalingGains <- TheColumns[str_detect(TheColumns, "Scaling")]
ScalingGains <- Luciernaga:::ColorPriority(ScalingGains)

VioletPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair3L)

BluePlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                      RepairVisits=Repair3L)

RedPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "Gain",
                     RepairVisits=Repair3L)

ScatterPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)

LaserDelayPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=LaserDelayGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)

ScalingPlotsGain <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScalingGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)
```

```{r}
x <- Gain_3L
TheColumns <- x %>% select(where(~is.numeric(.)||is.integer(.))) %>% colnames()
TheColumns <- setdiff(TheColumns, "TIME")
TheColumnNames <- TheColumns[str_detect(TheColumns, "rCV")]

VioletGains <- TheColumnNames[str_detect(TheColumnNames, "^V")]
BlueGains <- TheColumnNames[str_detect(TheColumnNames, "^B")]
RedGains <- TheColumnNames[str_detect(TheColumnNames, "^R")]

ScatterGains <- TheColumnNames[str_detect(TheColumnNames, "SC-")]
ScatterGains <- Luciernaga:::ScalePriority(ScatterGains)

VioletPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=VioletGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair3L)

BluePlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=BlueGains,
                      plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                      RepairVisits=Repair3L)

RedPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=RedGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = "%rCV",
                     RepairVisits=Repair3L)

ScatterPlotsRCV <- QC_Plots(x=x, FailedFlag=TRUE, MeasurementType=ScatterGains,
                     plotType = "individual", returntype = "plots", YAxisLabel = " ",
                     RepairVisits=Repair3L)
```

## {.sidebar}
Dashboard data for the **Cytek Aurora 3L** last updated on **`r TheDate`**

**Note:** Cytek QC Bead Lots were switched on 12-06-2024. MFI change on this date was expected.

**First Column: MFI** Median Fluorescent Intensity (MFI) values for QC beads acquired Before and After QC. Measures stability over time.
**Second Column: Gain** Gain (Voltage) values set for instrument after QC. Changes over time reflective of laser health.
**Third Column: rCV** Percentage Robust Coefficient of Variation (rCV) after QC. Higher values reflect decreased resolution between positive and negative for that detector.

For additional information concerning individual parameter tabs, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek Aurora cytometers, adapted from the [UMGCC FCSS InstrumentQC](https://umgccfcss.github.io/InstrumentQC/) project.

This dashboard was created with [Quarto](https://quarto.org/), based on work by [David Rach](https://github.com/DavidRach).

## MFI {.tabset}

```{r}
#| title: Violet
htmltools::tagList(lapply(VioletPlotsMFI, ggplotly))
```

```{r}
#| title: Blue
htmltools::tagList(lapply(BluePlotsMFI, ggplotly))
```

```{r}
#| title: Red
htmltools::tagList(lapply(RedPlotsMFI, ggplotly))
```

```{r}
#| title: Scatter
htmltools::tagList(lapply(ScatterPlotsMFI, ggplotly))
```

```{r}
#| title: LaserDelay
htmltools::tagList(lapply(LaserDelayPlotsGain, ggplotly))
```

```{r}
#| title: LaserScaling
htmltools::tagList(lapply(ScalingPlotsGain, ggplotly))
```

## Gain {.tabset}

```{r}
#| title: Violet
htmltools::tagList(lapply(VioletPlotsGain, ggplotly))
```

```{r}
#| title: Blue
htmltools::tagList(lapply(BluePlotsGain, ggplotly))
```

```{r}
#| title: Red
htmltools::tagList(lapply(RedPlotsGain, ggplotly))
```

```{r}
#| title: Scatter
htmltools::tagList(lapply(ScatterPlotsGain, ggplotly))
```

```{r}
#| title: LaserDelay
htmltools::tagList(lapply(LaserDelayPlotsGain, ggplotly))
```

```{r}
#| title: LaserScaling
htmltools::tagList(lapply(ScalingPlotsGain, ggplotly))
```

## rCV {.tabset}

```{r}
#| title: Violet
htmltools::tagList(lapply(VioletPlotsRCV, ggplotly))
```

```{r}
#| title: Blue
htmltools::tagList(lapply(BluePlotsRCV, ggplotly))
```

```{r}
#| title: Red
htmltools::tagList(lapply(RedPlotsRCV, ggplotly))
```

```{r}
#| title: Scatter
htmltools::tagList(lapply(ScatterPlotsRCV, ggplotly))
```

```{r}
#| title: LaserDelay
htmltools::tagList(lapply(LaserDelayPlotsGain, ggplotly))
```

```{r}
#| title: LaserScaling
htmltools::tagList(lapply(ScalingPlotsGain, ggplotly))
```
