---
format:
  dashboard:
    orientation: columns
aliases:
  - home.html
---
```{r}
#| message: FALSE
library(dplyr)
library(purrr)
library(stringr)
library(plotly)
library(Luciernaga)

Computer <- getwd()
MainFolder <- file.path(Computer, "data")
TheList <- c("3L", "5L")

# Updating Data
#walk(.x=TheList, MainFolder=MainFolder, .f=Luciernaga:::DailyQCParse)
```

```{r}
Gain_3L <- Luciernaga:::CurrentData(x="3L", MainFolder=MainFolder, type = "Gain")
Gain_5L <- Luciernaga:::CurrentData(x="5L", MainFolder=MainFolder, type = "Gain")
```

```{r}
#| eval: false
WindowOfInterest <- Sys.time() - months(24)

Gain_3L <- Gain_3L |> filter(DateTime >= WindowOfInterest)
Gain_5L <- Gain_5L |> filter(DateTime >= WindowOfInterest)
```

```{r}
# QC indicator detectors (per the Help page): FSC, SSC, and the "3" detectors
# (UV3/V3/B3/YG3/R3). The Gain archive names these without the "-A" area suffix,
# so we select them explicitly rather than relying on the default detectorType.
IndicatorDetectors <- "^FSC$|^SSC|V3$|B3$|R3$|YG3$"

The3L <- Luciernaga:::VisualQCSummary(x=Gain_3L, detectorType=IndicatorDetectors)
The5L <- Luciernaga:::VisualQCSummary(x=Gain_5L, detectorType=IndicatorDetectors)

Table3L <- Luciernaga:::SmallTable(data=The3L)
Table5L <- Luciernaga:::SmallTable(data=The5L)

# Cytek Aurora EVO (third instrument). Its processed Archive data may not be
# present yet (SpectroFlo 4.x parsing is pending), so build its table only when
# the Gain archive exists; otherwise leave it NULL for a placeholder.
EVOGainFile <- list.files(file.path(MainFolder, "EVO", "Archive"),
                          pattern = "Archived", full.names = TRUE)
HasEVO <- length(EVOGainFile) == 1 && nrow(read.csv(EVOGainFile[1], nrow = 1)) > 0
if (HasEVO) {
  Gain_EVO <- Luciernaga:::CurrentData(x="EVO", MainFolder=MainFolder, type = "Gain")
  TheEVO <- Luciernaga:::VisualQCSummary(x=Gain_EVO, detectorType=IndicatorDetectors)
  TableEVO <- Luciernaga:::SmallTable(data=TheEVO)
} else {
  TableEVO <- NULL
}
```

```{r}
#| include: false
#| echo: false

Computer <- getwd()
MainFolder <- file.path(Computer, "data")
TheName <- "HistoricalData.csv"
HistoricalPath <- file.path(MainFolder, TheName)
ShinyData <- read.csv(HistoricalPath, check.names = FALSE)
```

```{r}
x <- as.list(intersect(c("3L", "5L", "EVO"), unique(ShinyData$Instrument)))

DataForPlot <- Luciernaga:::QCHistoryArchive(x=x, historydata=ShinyData, timewindow=24)
Transposed <- t(DataForPlot)
colnames(Transposed) <- Transposed[1,]
Transposed <- Transposed[-1,]
DataForPlot1 <- data.frame(Transposed, check.names=FALSE)
DataForPlot1 <- DataForPlot1 %>% tibble::rownames_to_column(., var="Date")
DataForPlot1$Date <- as.Date(DataForPlot1$Date) 
Data <- DataForPlot1 |> arrange(desc(Date))

Data <- Data |> relocate("3L", "5L", .after="Date")

Data <- Data |> mutate(across(-Date, ~ na_if(., "Unknown")))

AltData <- Data |> dplyr::filter(Date > lubridate::ymd("2023-04-10"))

GlobalSummary <- Luciernaga:::SmallTableGlobal(AltData)
```

```{r}
TheDate <- Data |> slice(1) |> pull(Date)
```

```{r}
TheStatus3L <- Luciernaga:::CurrentStatus(x="3L", data=Data) %>% Luciernaga:::InstrumentText(.)
TheColor3L <- Luciernaga:::CurrentStatus(x="3L", data=Data) %>% Luciernaga:::InstrumentColor(.)
TheStatus5L <- Luciernaga:::CurrentStatus(x="5L", data=Data) %>% Luciernaga:::InstrumentText(.)
TheColor5L <- Luciernaga:::CurrentStatus(x="5L", data=Data) %>% Luciernaga:::InstrumentColor(.)

# Cytek Aurora EVO status: real status once it appears in HistoricalData,
# otherwise an "Awaiting data" placeholder so the home page renders cleanly.
if ("EVO" %in% names(Data)) {
  TheStatusEVO <- Luciernaga:::CurrentStatus(x="EVO", data=Data) %>% Luciernaga:::InstrumentText(.)
  TheColorEVO  <- Luciernaga:::CurrentStatus(x="EVO", data=Data) %>% Luciernaga:::InstrumentColor(.)
} else {
  TheStatusEVO <- "Awaiting data"
  TheColorEVO  <- "secondary"
}
```

## {.sidebar}
Dashboard data last updated on **`r TheDate`**

**Definitions:**

**Pass:** All gains within 100% baseline and all RCVs <6% for all detectors.

**Caution:** All gains within 100% baseline, but at least one detector had a RCV above the >6% cutoff. Instrument remains usable but resolution for fluorophores on the failed detector may decrease. 

**Fail:** Either a gain exceeded 100% baseline, or RCVs exceeded >6% for at least one indicator detector. Significant variation and batch effects may occcur. 

For additional information, navigate to the [Help](help.qmd) page.

**About**

This dashboard contains the visualized QC data for the Cytek cytometers at the Sydney Cytometry Core Facility

For additional flow cytometry resources, visit our [website](https://sydneycytometry.org.au/)

This dashboard was created with [Quarto](https://quarto.org/) and is maintained by [David Rach](https://github.com/DavidRach)

## First {width="30%"}

### Row {height="33%"}

#### Column

```{r}
#| content: valuebox
#| title: "3L"
#| icon: cup-hot

list(value = paste0("QC Status: ", TheStatus3L),
     color = TheColor3L)

```

#### Column


### Row  {height="33%"}

#### Column

```{r}
#| content: valuebox
#| title: "5L"
#| icon: cup-hot

list(value = paste0("QC Status: ", TheStatus5L),
     color = TheColor5L)
```

#### Column {scrolling=FALSE}

### Row {height="34%"}

#### Column

```{r}
#| content: valuebox
#| title: "EVO"
#| icon: cup-hot

list(value = paste0("QC Status: ", TheStatusEVO),
     color = TheColorEVO)
```

#### Column

## Second {.tabset}
```{r}
#| title: 3L
Table3L
```


```{r}
#| title: 5L
Table5L
```


```{r}
#| title: EVO
if (!is.null(TableEVO)) TableEVO else htmltools::HTML("<em>Awaiting first QC data</em>")
```


## Third {.tabset}{width="40%"}
```{r}
#| title: Instruments
GlobalSummary
```